---
title: "ASESII 24A02 Project-Based Quiz 1"
subtitle: "Ridge, Lasso, and Regularization for Neural Features"
author: "Group XX: [names and student IDs]"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    fig_crop: false
  html_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
link-citations: true
params:
  group_number: 2
fontsize: 11pt
geometry: margin=1in
---

<!--
Copy this template into your group submission folder, rename it
GroupXX_ProjectQuiz1.Rmd, place references.bib beside it, and set the group
number parameter before rendering. Remove instructional comments from the
final report.
-->

# Authorship and contribution statement {-}

We confirm that every group member can explain the submitted analysis. AI use is reported in `GroupXX_AI_Log.csv`, and the group checked every AI-assisted item used in this report.

| Student | ID | Main contributions | Verification performed |
|---|---|---|---|
| [Name] | [ID] | [Work] | [Check] |

# Abstract {-}

*Maximum 150 words: prediction problem, methods, main evidence, and recommendation.*

# Reproducible setup {-}

```{r setup, message=FALSE, warning=FALSE}
required_packages <- c("tidyverse", "glmnet", "broom", "knitr")
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop("Install the required packages before rendering: ",
       paste(missing_packages, collapse = ", "))
}

library(tidyverse)
library(glmnet)
library(broom)
library(knitr)

group_number <- as.integer(params$group_number)
if (length(group_number) != 1L || is.na(group_number) || group_number < 1L) {
  stop("Set params$group_number to your assigned positive group number.")
}

seeds <- list(
  split = 240200L + group_number,
  cv = 240300L + group_number,
  features = 240400L + group_number
)
```

## Shared helper functions {-}

```{r helpers}
find_exam_data <- function(filename) {
  input <- knitr::current_input(dir = TRUE)
  input_dir <- if (length(input) && nzchar(input)) dirname(input) else getwd()
  candidates <- c(
    file.path(input_dir, "data", filename),
    file.path(input_dir, "..", "data", filename),
    file.path(getwd(), "data", filename)
  )
  existing <- candidates[file.exists(candidates)]
  if (!length(existing)) stop("Cannot locate ", filename, " by a relative path.")
  hits <- unique(normalizePath(existing, winslash = "/", mustWork = TRUE))
  if (length(hits) > 1L && length(unique(unname(tools::md5sum(hits)))) > 1L) {
    stop("Multiple nonidentical copies of ", filename, " were found.")
  }
  hits[[1L]]
}

split_rows <- function(n, train_prop = 0.80, seed) {
  stopifnot(n >= 10L, train_prop > 0, train_prop < 1)
  set.seed(seed)
  train <- sort(sample.int(n, floor(train_prop * n), replace = FALSE))
  list(train = train, test = setdiff(seq_len(n), train))
}

fit_scaler <- function(x, tol = 1e-12) {
  x <- as.matrix(x)
  center <- colMeans(x)
  centered <- sweep(x, 2L, center, "-")
  scale <- sqrt(colMeans(centered^2))
  keep <- is.finite(scale) & scale > tol
  if (!any(keep)) stop("No nonconstant training predictor remains.")
  list(
    center = center[keep],
    scale = scale[keep],
    keep = keep,
    dropped = colnames(x)[!keep]
  )
}

apply_scaler <- function(x, prep) {
  x <- as.matrix(x)[, prep$keep, drop = FALSE]
  x <- sweep(x, 2L, prep$center, "-")
  sweep(x, 2L, prep$scale, "/")
}

make_foldid <- function(n, k = 5L, seed) {
  if (k < 3L || k > n) stop("Require 3 <= k <= number of training rows.")
  set.seed(seed)
  sample(rep(seq_len(k), length.out = n))
}

fit_cv_glmnet <- function(x, y, alpha, foldid) {
  glmnet::cv.glmnet(
    x = x,
    y = y,
    family = "gaussian",
    alpha = alpha,
    foldid = foldid,
    standardize = FALSE,
    intercept = TRUE,
    type.measure = "mse"
  )
}

score_regression <- function(y, prediction) {
  y <- as.numeric(y)
  prediction <- as.numeric(prediction)
  if (length(y) != length(prediction) || any(!is.finite(c(y, prediction)))) {
    stop("Metrics require equal-length finite vectors.")
  }
  c(RMSE = sqrt(mean((y - prediction)^2)),
    MAE = mean(abs(y - prediction)))
}

count_nonzero <- function(fit, s = "lambda.min", tol = 1e-8) {
  b <- as.matrix(stats::coef(fit, s = s))
  slopes <- b[rownames(b) != "(Intercept)", 1L]
  sum(abs(slopes) > tol)
}

safe_condition_numbers <- function(x, lambda, tol = 1e-10) {
  eigenvalues <- eigen(
    crossprod(x) / nrow(x),
    symmetric = TRUE,
    only.values = TRUE
  )$values
  largest <- max(eigenvalues)
  smallest <- max(min(eigenvalues), 0)
  cutoff <- tol * max(1, largest)
  c(
    gram = if (smallest <= cutoff) Inf else largest / smallest,
    ridge = (largest + lambda) / (smallest + lambda)
  )
}
```

# Problem 1. Prediction design and leakage control

## 1A. Target and predictors

```{r data-configuration, eval=FALSE}
config <- if (group_number %% 2L == 1L) {
  list(
    file = "fat.csv",
    response = "brozek",
    excluded = c("brozek", "siri", "density", "free")
  )
} else {
  list(
    file = "prostate.csv",
    response = "lpsa",
    excluded = "lpsa"
  )
}

data_raw <- read.csv(find_exam_data(config$file), check.names = FALSE)
predictors <- setdiff(names(data_raw), config$excluded)
analysis_data <- data_raw[, c(config$response, predictors), drop = FALSE]

if (anyNA(analysis_data)) stop("Declare a training-only missing-value plan.")
if (!all(vapply(analysis_data, is.numeric, logical(1)))) {
  stop("The assigned response and predictors must be numeric.")
}
```

*Define the prediction task and justify every exclusion.*

## 1B. Split and train-only preprocessing

```{r split-and-scale}

add_nonlinear_features <- function(x) {
  x <- as.matrix(x)
  p <- ncol(x)
  n_rows <- nrow(x)
  col_names <- colnames(x)
  if (is.null(col_names)) {
    col_names <- paste0("V", seq_len(p))
  }
  
  # 1. Calculate the number of non-linear terms
  # p squared terms + p*(p-1)/2 interaction terms
  num_nonlinear <- p + (p * (p - 1)) / 2
  
  # 2. Pre-allocate the non-linear matrix for performance
  non_linear_mat <- matrix(NA, nrow = n_rows, ncol = num_nonlinear)
  non_linear_names <- character(num_nonlinear)
  
  col_idx <- 1
  
  # 3. Generate squared terms (x_i ^ 2)
  for (i in seq_len(p)) {
    non_linear_mat[, col_idx] <- x[, i]^2
    non_linear_names[col_idx] <- paste0(col_names[i], "_sq")
    col_idx <- col_idx + 1
  }
  
  # 4. Generate cross-interaction terms (x_i * x_j)
  if (p > 1) {
    for (i in seq_len(p - 1)) {
      for (j in (i + 1):p) {
        non_linear_mat[, col_idx] <- x[, i] * x[, j]
        non_linear_names[col_idx] <- paste0(col_names[i], "_x_", col_names[j])
        col_idx <- col_idx + 1
      }
    }
  }
  
  # Assign column names to the non-linear matrix
  colnames(non_linear_mat) <- non_linear_names
  
  # 5. Combine the original linear matrix and the new non-linear matrix
  cbind(x, non_linear_mat)
}

# Split once using the declared group-specific seed.
split <- split_rows(
    n = nrow(analysis_data),
    train_prop = 0.80,
    seed = seeds$split
)

train_id <- split$train
test_id <- split$test

stopifnot(
    length(intersect(train_id, test_id)) == 0L
)

x_raw <- as.matrix(analysis_data[, predictors, drop = FALSE])
y_raw <- analysis_data[[config$response]]

x_raw_nonlinear <- add_nonlinear_features(x_raw)

x_prep <- fit_scaler(x_raw_nonlinear[train_id, , drop = FALSE])
x_raw_nonlinear_scaled <- apply_scaler(x_raw_nonlinear, x_prep)

x_prep <- fit_scaler(x_raw[train_id, , drop = FALSE])
x_raw_scaled <- apply_scaler(x_raw, x_prep)

#official data

#non linear data
x_train_nonlinear <- x_raw_nonlinear_scaled[train_id, , drop = FALSE]
x_test_nonlinear <- x_raw_nonlinear_scaled[test_id, , drop = FALSE]
y_train_nonlinear <- y_raw[train_id]
y_test_nonlinear <- y_raw[test_id]

#linear data
x_train <- x_raw_scaled[train_id, , drop = FALSE]
x_test <- x_raw_scaled[test_id, , drop = FALSE]
y_train <- y_raw[train_id]
y_test <- y_raw[test_id]

# Create one shared five-fold assignment for all training comparisons.
foldid <- make_foldid(
    n = nrow(x_train),
    k = 5L,
    seed = seeds$cv
)


```

*Report seeds, dimensions, scaling, folds, and leakage safeguards.*

## 1C. Training-data exploration

```{r training-eda, eval=FALSE}
# Use analysis_data[train_id, ] only. Add focused plots and summaries.
```

*Interpret multicollinearity, one anomaly, and the question regularization will address.*

# Problem 2. OLS, ridge, and lasso

## 2A. OLS baseline

```{r ols, eval=FALSE}
# Fit OLS on x_train and y_train. Do not access y_test here.

ols_stepwise_regression <- function(x_train,
                        y_train,
                        direction = "both") {
  df = as.data.frame(x_train)
  df$y = y_train
  full_model <- lm(y ~ ., data = df)

  best_model <- step(
    full_model,
    direction = direction,
    trace = 1
  )

  return(best_model)
}

ols_stepwise_regression(x_train_nonlinear, y_train_nonlinear)


```

## 2B. Ridge

```{r ridge, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 0, foldid = foldid).
# Report lambda.min, lambda.1se, paths, coefficients, and CV evidence.
```

## 2C. Lasso

```{r lasso, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 1, foldid = foldid).
# Report lambda.min, lambda.1se, paths, and nonzero predictors.
```

## 2D. Training-based comparison and locked core model

```{r core-comparison, eval=FALSE}
# Build a comparison from training/CV quantities only.
```

**Core model nominated before holdout evaluation:** [method and reason]

## 2E. Perturbation or stability check

**Prediction before running the check:** [prediction]

```{r perturbation, eval=FALSE}
# Implement one allowed training-only stability check.
```

**Observed result and explanation:** [result]

# Problem 3. Mathematical mechanisms

## 3A. Ridge and conditioning

*Write the ridge derivation. Define dimensions and explain uniqueness.*

```{r ridge-conditioning, eval=FALSE}
# Use x_train and the fitted ridge lambda.min.
# safe_condition_numbers(x_train, lambda = ...)
```

## 3B. Lasso optimality and soft thresholding

*Derive the zero/nonzero optimality conditions and the orthonormal-design formula.*

## 3C. Connect algebra to evidence

*Connect one specific numerical result to Problem 3A or 3B.*

# Problem 4. Elastic net and a neural-feature bridge

## 4A. Research question and source map

**Research direction:** [weight decay / sparse weights / pruning / dropout]

| Claim | Exact primary source and locator | What it establishes | What it does not establish |
|---|---|---|---|
| [Claim] | [Citation, section/page/equation] | [Support] | [Boundary] |

## 4B. Elastic net on original predictors

```{r original-elastic-net, eval=FALSE}
alpha_grid <- c(0.10, 0.25, 0.50, 0.75, 0.90)

# Fit one cv.glmnet object per alpha with the same foldid.
# Select alpha and lambda using training CV only.

fit_elastic_net_4B <- function(x, y, foldid, alpha_list){
  fits <- list()
  for (alpha in alpha_list){
        fits[[as.character((alpha))]] <- cv.glmnet(
        x = x,
        y = y,
        alpha = alpha,
        foldid = foldid
    )
  }
  return(fits)
}

plot_coef_path_4B <- function(glmnet_fit, alpha, title) {

  if (!is.null(alpha)) {
    title <- paste0(title, " (alpha = ", alpha, ")")
  } 
  
  coef_mat <- as.matrix(glmnet_fit$beta)
  colnames(coef_mat) <- glmnet_fit$lambda

  coef_df <- as.data.frame(coef_mat) |>
    tibble::rownames_to_column("variable") |>
    tidyr::pivot_longer(
      -variable,
      names_to = "lambda",
      values_to = "coefficient"
    ) |>
    dplyr::mutate(
      lambda = as.numeric(lambda)
    )

  ggplot2::ggplot(
    coef_df,
    ggplot2::aes(
      x = lambda,
      y = coefficient,
      colour = variable
    )
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = title,
      x = expression(lambda),
      y = "Coefficient"
    ) +
    ggplot2::theme_minimal()
}

plot_nonzero_slopes_4B <- function(glmnet_fit, alpha, title) {
  
  if (!is.null(alpha)) {
    title <- paste0(title, " (alpha = ", alpha, ")")
  } 

  df_data <- tibble::tibble(
    lambda = glmnet_fit$lambda,
    df = glmnet_fit$df
  )

  ggplot2::ggplot(
    df_data,
    ggplot2::aes(
      x = lambda,
      y = df,
    )
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = title,
      x = expression(lambda),
      y = "number of none-zero slopes"
    ) +
    ggplot2::theme_minimal()
}

plot_cross_validation_error <- function(cv_glmnet_fit, alpha, title) {
  
  if (!is.null(alpha)) {
    title <- paste0(title, " (alpha = ", alpha, ")")
  } 

  error_data <- tibble::tibble(
    lambda = cv_glmnet_fit$lambda,
    error = cv_glmnet_fit$cvm
  )

  ggplot2::ggplot(
    error_data,
    ggplot2::aes(
      x = lambda,
      y = error,
    )
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = title,
      x = expression(lambda),
      y = "cross-validation error"
    ) +
    ggplot2::theme_minimal()
}

elastic_net_list <- fit_elastic_net_4B(x_train, y_train, foldid, alpha_grid)

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]$glmnet.fit
  print(plot_coef_path_4B(fit, alpha, "Elastic Net Coefficient Paths"))
}

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]$glmnet.fit
  print(plot_nonzero_slopes_4B(fit, alpha, "Number of Non-Zero Slopes"))
}

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]
  print(plot_cross_validation_error(fit, alpha, "Cross-Validation Error"))
}

cv_results <- tibble::tibble(
  alpha = numeric(),
  lambda = numeric(),
  cv_error = numeric()
)

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]
  idx <- which(fit$lambda == fit$lambda.min)
  error <- fit$cvm[idx]
  cv_results <- dplyr::add_row(
    cv_results,
    alpha = as.numeric(alpha),
    lambda = fit$lambda.min,
    cv_error = error
  )
}

best_elastic_net_param <- cv_results |> dplyr::slice_min(cv_error, n=1)
best_elastic_net_param <- best_elastic_net_param[1, ]
best_elastic_net_model <- elastic_net_list[[as.character(
  best_elastic_net_param$alpha
)]]
```



## 4C. Fixed ReLU features

```{r random-relu-features, eval=FALSE}
relu <- function(z, A, C){
  z <- z %*% t(A)
  C <- matrix(C, nrow = nrow(z), ncol = length(C), byrow = TRUE)
  z <- z + C
  return (pmax(z, 0))
} 
M <- 200L
p <- ncol(x_train)

# TODO:
# 1. Generate A and bias once from seeds$features using the stated distributions.
# 2. Apply the same A and bias to x_train and x_test.
# 3. Fit a scaler to H_train only and apply it to H_test.
# 4. Fit ridge, lasso, and elastic net on H_train with the shared foldid.

set.seed(22)

A <- matrix(rnorm(M * p, mean = 0, sd = sqrt(1/p)),
            nrow = M, ncol = p)

C <- matrix(rnorm(M * 1, mean = 0, sd = 0.5 ),
            nrow = M, ncol = 1)

x_hidden <- relu(x_train, A, C)
same_col <- apply(x_hidden, 2, function(col) all(col == col[1]))
print(same_col)

cv_ridge <- cv.glmnet(
  x = x_hidden,
  y = y_train,
  alpha = 0,     
  foldid = foldid 
)

cv_lasso <- cv.glmnet(
  x = x_hidden,
  y = y_train,
  alpha = 1,     
  foldid = foldid 
)

cv_elastic_net_list <- fit_elastic_net_4B(x_hidden, y_train, foldid, alpha_grid)

plot_cross_validation_error(cv_ridge, NULL, "Cross Validation Error of Ridge")
plot_nonzero_slopes_4B(cv_ridge$glmnet.fit, NULL, "Number of Active Hidden Features of Ridge")

plot_cross_validation_error(cv_lasso, NULL, "Cross Validation Error of Lasso")
plot_nonzero_slopes_4B(cv_lasso$glmnet.fit, NULL, "Number of Active Hidden Features of Lasso")

for (alpha in names(cv_elastic_net_list)) {
  fit <- cv_elastic_net_list[[alpha]]$glmnet.fit
  print(plot_nonzero_slopes_4B(fit, alpha, "Number of Non-Zero Slopes of Elastic Net"))
}

for (alpha in names(cv_elastic_net_list)) {
  fit <- cv_elastic_net_list[[alpha]]
  print(plot_cross_validation_error(fit, alpha, "Cross-Validation Error of Elastic Net"))
}

cv_results <- tibble::tibble(
  alpha = numeric(),
  lambda = numeric(),
  cv_error = numeric()
)

for (alpha in names(cv_elastic_net_list)) {
  fit <- cv_elastic_net_list[[alpha]]
  idx <- which(fit$lambda == fit$lambda.min)
  error <- fit$cvm[idx]
  cv_results <- dplyr::add_row(
    cv_results,
    alpha = as.numeric(alpha),
    lambda = fit$lambda.min,
    cv_error = error
  )
}

best_elastic_net_param_4C <- cv_results |> dplyr::slice_min(cv_error, n=1)
best_elastic_net_param_4C <- best_elastic_net_param_4C[1, ]
best_elastic_net_model_4C <- cv_elastic_net_list[[as.character(
  best_elastic_net_param_4C$alpha
)]]$glmnet.fit

best_ridge_model_4C <- cv_ridge
best_lasso_model_4C <- cv_lasso

```

*State which weights are fixed, which are learned, and how active features are counted.*

## 4D. Locked declaration and final holdout evaluation

> **Locked before viewing `y_test`:** [preprocessing, candidate specifications, selected tuning values, feature seed, nominated deployment model, metrics]

```{r final-holdout-evaluation, eval=FALSE}
# This must be the first analysis chunk that uses y_test for evaluation.
# Generate all fixed candidate predictions and one RMSE/MAE table here.

MAE <- function(actual, predicted) {
  mean(abs(actual - predicted))
}

RMSE <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}


```

*State whether the holdout evidence supports the predeclared choice. Do not retune.*

## 4E. Deep-learning connection and limitations

*Distinguish penalty, weight decay, output-weight sparsity, pruning, dropout, and a trained deep network. Cite the sources from 4A and discuss at least two limitations.*

# Problem 5. Report quality and reproducibility

## 5A. Reproduction record

```text
1. Open a clean R session.
2. [Exact commands and folder layout.]
3. Render GroupXX_ProjectQuiz1.Rmd.
```

## 5B. Recommendation and limitations

*Give the final recommendation, distinguish prediction from interpretation, and state limitations.*

## 5C. AI verification record

| AI-assisted item | How it was checked | Evidence | Modification or rejection |
|---|---|---|---|
| [Item] | [Method] | [Test/source/output] | [Decision] |

# Preflight and session information {-}

```{r preflight, eval=FALSE}
# Add every object required to reproduce the final table and figures.
required_objects <- c(
  "train_id", "test_id", "x_train", "x_test", "y_train", "y_test", "foldid"
)
missing_objects <- setdiff(required_objects, ls())
if (length(missing_objects)) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}
stopifnot(length(intersect(train_id, test_id)) == 0L)
```

```{r session-info}
sessionInfo()
```
