---
title: "ASESII 24A02 Project-Based Quiz 1"
subtitle: "Ridge, Lasso, and Regularization for Neural Features"
author: "Group XX: [names and student IDs]"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    fig_crop: false
  html_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
link-citations: true
params:
  group_number: 2
fontsize: 11pt
geometry: margin=1in
---

<!--
Copy this template into your group submission folder, rename it
GroupXX_ProjectQuiz1.Rmd, place references.bib beside it, and set the group
number parameter before rendering. Remove instructional comments from the
final report.
-->

# Authorship and contribution statement {-}

We confirm that every group member can explain the submitted analysis. AI use is reported in `GroupXX_AI_Log.csv`, and the group checked every AI-assisted item used in this report.

| Student | ID | Main contributions | Verification performed |
|---|---|---|---|
| [Name] | [ID] | [Work] | [Check] |

# Abstract {-}

*Maximum 150 words: prediction problem, methods, main evidence, and recommendation.*

# Reproducible setup {-}

```{r setup, message=FALSE, warning=FALSE}
required_packages <- c("tidyverse", "glmnet", "broom", "knitr")
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop("Install the required packages before rendering: ",
       paste(missing_packages, collapse = ", "))
}

library(tidyverse)
library(glmnet)
library(broom)
library(knitr)

group_number <- as.integer(params$group_number)
if (length(group_number) != 1L || is.na(group_number) || group_number < 1L) {
  stop("Set params$group_number to your assigned positive group number.")
}

seeds <- list(
  split = 240200L + group_number,
  cv = 240300L + group_number,
  features = 240400L + group_number
)
```

## Shared helper functions {-}

```{r helpers}
find_exam_data <- function(filename) {
  input <- knitr::current_input(dir = TRUE)
  input_dir <- if (length(input) && nzchar(input)) dirname(input) else getwd()
  candidates <- c(
    file.path(input_dir, "data", filename),
    file.path(input_dir, "..", "data", filename),
    file.path(getwd(), "data", filename)
  )
  existing <- candidates[file.exists(candidates)]
  if (!length(existing)) stop("Cannot locate ", filename, " by a relative path.")
  hits <- unique(normalizePath(existing, winslash = "/", mustWork = TRUE))
  if (length(hits) > 1L && length(unique(unname(tools::md5sum(hits)))) > 1L) {
    stop("Multiple nonidentical copies of ", filename, " were found.")
  }
  hits[[1L]]
}

split_rows <- function(n, train_prop = 0.80, seed) {
  stopifnot(n >= 10L, train_prop > 0, train_prop < 1)
  set.seed(seed)
  train <- sort(sample.int(n, floor(train_prop * n), replace = FALSE))
  list(train = train, test = setdiff(seq_len(n), train))
}

fit_scaler <- function(x, tol = 1e-12) {
  x <- as.matrix(x)
  center <- colMeans(x)
  centered <- sweep(x, 2L, center, "-")
  scale <- sqrt(colMeans(centered^2))
  keep <- is.finite(scale) & scale > tol
  if (!any(keep)) stop("No nonconstant training predictor remains.")
  list(
    center = center[keep],
    scale = scale[keep],
    keep = keep,
    dropped = colnames(x)[!keep]
  )
}

apply_scaler <- function(x, prep) {
  x <- as.matrix(x)[, prep$keep, drop = FALSE]
  x <- sweep(x, 2L, prep$center, "-")
  sweep(x, 2L, prep$scale, "/")
}

make_foldid <- function(n, k = 5L, seed) {
  if (k < 3L || k > n) stop("Require 3 <= k <= number of training rows.")
  set.seed(seed)
  sample(rep(seq_len(k), length.out = n))
}

fit_cv_glmnet <- function(x, y, alpha, foldid) {
  glmnet::cv.glmnet(
    x = x,
    y = y,
    family = "gaussian",
    alpha = alpha,
    foldid = foldid,
    standardize = FALSE,
    intercept = TRUE,
    type.measure = "mse"
  )
}

score_regression <- function(y, prediction) {
  y <- as.numeric(y)
  prediction <- as.numeric(prediction)
  if (length(y) != length(prediction) || any(!is.finite(c(y, prediction)))) {
    stop("Metrics require equal-length finite vectors.")
  }
  c(RMSE = sqrt(mean((y - prediction)^2)),
    MAE = mean(abs(y - prediction)))
}

count_nonzero <- function(fit, s = "lambda.min", tol = 1e-8) {
  b <- as.matrix(stats::coef(fit, s = s))
  slopes <- b[rownames(b) != "(Intercept)", 1L]
  sum(abs(slopes) > tol)
}

safe_condition_numbers <- function(x, lambda, tol = 1e-10) {
  eigenvalues <- eigen(
    crossprod(x) / nrow(x),
    symmetric = TRUE,
    only.values = TRUE
  )$values
  largest <- max(eigenvalues)
  smallest <- max(min(eigenvalues), 0)
  cutoff <- tol * max(1, largest)
  c(
    gram = if (smallest <= cutoff) Inf else largest / smallest,
    ridge = (largest + lambda) / (smallest + lambda)
  )
}
```

# Problem 1. Prediction design and leakage control

## 1A. Target and predictors

```{r data-configuration, eval=FALSE}
config <- if (group_number %% 2L == 1L) {
  list(
    file = "fat.csv",
    response = "brozek",
    excluded = c("brozek", "siri", "density", "free")
  )
} else {
  list(
    file = "prostate.csv",
    response = "lpsa",
    excluded = "lpsa"
  )
}

data_raw <- read.csv(find_exam_data(config$file), check.names = FALSE)
predictors <- setdiff(names(data_raw), config$excluded)
analysis_data <- data_raw[, c(config$response, predictors), drop = FALSE]

if (anyNA(analysis_data)) stop("Declare a training-only missing-value plan.")
if (!all(vapply(analysis_data, is.numeric, logical(1)))) {
  stop("The assigned response and predictors must be numeric.")
}
```

*Define the prediction task and justify every exclusion.*

## 1B. Split and train-only preprocessing

```{r split-and-scale}

add_nonlinear_features <- function(x) {
  x <- as.matrix(x)
  p <- ncol(x)
  n_rows <- nrow(x)
  col_names <- colnames(x)
  if (is.null(col_names)) {
    col_names <- paste0("V", seq_len(p))
  }
  
  # 1. Calculate the number of non-linear terms
  # p squared terms + p*(p-1)/2 interaction terms
  num_nonlinear <- p + (p * (p - 1)) / 2
  
  # 2. Pre-allocate the non-linear matrix for performance
  non_linear_mat <- matrix(NA, nrow = n_rows, ncol = num_nonlinear)
  non_linear_names <- character(num_nonlinear)
  
  col_idx <- 1
  
  # 3. Generate squared terms (x_i ^ 2)
  for (i in seq_len(p)) {
    non_linear_mat[, col_idx] <- x[, i]^2
    non_linear_names[col_idx] <- paste0(col_names[i], "_sq")
    col_idx <- col_idx + 1
  }
  
  # 4. Generate cross-interaction terms (x_i * x_j)
  if (p > 1) {
    for (i in seq_len(p - 1)) {
      for (j in (i + 1):p) {
        non_linear_mat[, col_idx] <- x[, i] * x[, j]
        non_linear_names[col_idx] <- paste0(col_names[i], "_x_", col_names[j])
        col_idx <- col_idx + 1
      }
    }
  }
  
  # Assign column names to the non-linear matrix
  colnames(non_linear_mat) <- non_linear_names
  
  # 5. Combine the original linear matrix and the new non-linear matrix
  cbind(x, non_linear_mat)
}

# Split once using the declared group-specific seed.
split <- split_rows(
    n = nrow(analysis_data),
    train_prop = 0.80,
    seed = seeds$split
)

train_id <- split$train
test_id <- split$test

stopifnot(
    length(intersect(train_id, test_id)) == 0L
)

x_raw <- as.matrix(analysis_data[, predictors, drop = FALSE])
y_raw <- analysis_data[[config$response]]

x_raw_nonlinear <- add_nonlinear_features(x_raw)

x_prep <- fit_scaler(x_raw_nonlinear[train_id, , drop = FALSE])
x_raw_nonlinear_scaled <- apply_scaler(x_raw_nonlinear, x_prep)

x_prep <- fit_scaler(x_raw[train_id, , drop = FALSE])
x_raw_scaled <- apply_scaler(x_raw, x_prep)

#official data

#non linear data
x_train_nonlinear <- x_raw_nonlinear_scaled[train_id, , drop = FALSE]
x_test_nonlinear <- x_raw_nonlinear_scaled[test_id, , drop = FALSE]
y_train_nonlinear <- y_raw[train_id]
y_test_nonlinear <- y_raw[test_id]

#linear data
x_train <- x_raw_scaled[train_id, , drop = FALSE]
x_test <- x_raw_scaled[test_id, , drop = FALSE]
y_train <- y_raw[train_id]
y_test <- y_raw[test_id]

# Create one shared five-fold assignment for all training comparisons.
foldid <- make_foldid(
    n = nrow(x_train),
    k = 5L,
    seed = seeds$cv
)


```

*Report seeds, dimensions, scaling, folds, and leakage safeguards.*

## 1C. Training-data exploration

```{r training-eda, eval=FALSE}
# Use analysis_data[train_id, ] only. Add focused plots and summaries.
```

*Interpret multicollinearity, one anomaly, and the question regularization will address.*

# Problem 2. OLS, ridge, and lasso

## 2A. OLS baseline

```{r ols, eval=FALSE}
# Fit OLS on x_train and y_train. Do not access y_test here.

ols_stepwise_regression <- function(x_train,
                        y_train,
                        direction = "both") {
  df = as.data.frame(x_train)
  df$y = y_train
  full_model <- lm(y ~ ., data = df)

  best_model <- step(
    full_model,
    direction = direction,
    trace = 1
  )

  return(best_model)
}

ols_stepwise_regression(x_train_nonlinear, y_train_nonlinear)


```

## 2B. Ridge

```{r ridge, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 0, foldid = foldid).
# Report lambda.min, lambda.1se, paths, coefficients, and CV evidence.
```

## 2C. Lasso

```{r lasso, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 1, foldid = foldid).
# Report lambda.min, lambda.1se, paths, and nonzero predictors.
```

## 2D. Training-based comparison and locked core model

```{r core-comparison, eval=FALSE}
# Build a comparison from training/CV quantities only.
```

**Core model nominated before holdout evaluation:** [method and reason]

## 2E. Perturbation or stability check

**Prediction before running the check:** [prediction]

```{r perturbation, eval=FALSE}
# Implement one allowed training-only stability check.
```

**Observed result and explanation:** [result]

# Problem 3. Mathematical mechanisms

## 3A. Ridge and conditioning

*Write the ridge derivation. Define dimensions and explain uniqueness.*

```{r ridge-conditioning, eval=FALSE}
# Use x_train and the fitted ridge lambda.min.
# safe_condition_numbers(x_train, lambda = ...)
```

## 3B. Lasso optimality and soft thresholding

*Derive the zero/nonzero optimality conditions and the orthonormal-design formula.*

## 3C. Connect algebra to evidence

*Connect one specific numerical result to Problem 3A or 3B.*

# Problem 4. Elastic net and a neural-feature bridge

## 4A. Research Question and Source Map

**Research direction:** $L_2$ penalties and weight decay.

**Research question:** Under what optimization conditions are $L_2$ penalties equivalent to direct weight decay, and through what mechanisms can weight decay improve generalization in neural networks?

### $L_2$ regularization and direct weight decay

$L_2$ regularization is usually treated as equivalent to weight decay because both reduce the magnitude of model parameters. However, there is a difference in their nature, so this equivalence does not always hold.

Let $f_t(\boldsymbol{\theta})$ denote the loss on the mini-batch used at iteration $t$. Adding an $L_2$ penalty gives the regularized objective

$$
f_t^{\mathrm{reg}}(\boldsymbol{\theta})
=
f_t(\boldsymbol{\theta})
+
\frac{\lambda_{L2}}{2}
\left\|
\boldsymbol{\theta}
\right\|_2^2.
$$

Under standard stochastic gradient (SGD) descent with learning rate $\alpha$, the resulting update is

$$
\begin{aligned}
\boldsymbol{\theta}_{t+1}
&=
\left(
1-\alpha\lambda_{L2}
\right)
\boldsymbol{\theta}_t
-
\alpha
\nabla f_t(\boldsymbol{\theta}_t).
\end{aligned}
$$

Direct weight decay instead shrinks the parameters separately from the loss-gradient update:

$$
\boldsymbol{\theta}_{t+1}
=
\left(
1-\lambda_{WD}
\right)
\boldsymbol{\theta}_t
-
\alpha
\nabla f_t(\boldsymbol{\theta}_t).
$$

Therefore, the two procedures generate identical SGD updates when

$$
\lambda_{WD}
=
\alpha\lambda_{L2}.
$$


Thus, with standard SGD, $L_2$ regularization and weight decay are equivalent only when the $L_2$ coefficient is appropriately scaled by the learning rate [@loshchilov2019].

### Why the equivalence fails for adaptive optimizers

The equivalence generally fails when different parameters are scaled differently. Let $M_t$ denote the preconditioning matrix used by an adaptive optimizer. Applying an $L_2$ penalty gives

$$
\boldsymbol{\theta}_{t+1}
=
\boldsymbol{\theta}_t
-
\alpha M_t
\nabla f_t(\boldsymbol{\theta}_t)
-
\alpha\lambda_{L2}
M_t\boldsymbol{\theta}_t.
$$

In contrast, decoupled weight decay gives

$$
\boldsymbol{\theta}_{t+1}
=
\left(
1-\lambda_{WD}
\right)
\boldsymbol{\theta}_t
-
\alpha M_t
\nabla f_t(\boldsymbol{\theta}_t).
$$

With $L_2$ regularization, the penalty gradient is transformed by the same adaptive preconditioner as the data-loss gradient, while the direct weight decay applies the shrinkage directly and separately. Unless $M_t = kI$ for some $k \in \mathbb{R}$, no single $L_2$ coefficient can make the two updates equivalent for all parameters [@loshchilov2019].

This distinction motivated AdamW, which separates weight decay from Adam's adaptive gradient calculation. In the image-classification experiments reported by Loshchilov and Hutter, AdamW produced better generalization than Adam with an ordinary $L_2$ penalty and also made the learning-rate and weight-decay hyperparameters easier to tune independently [@loshchilov2019].

### How weight decay can regularize neural networks

Zhang et al. investigated weight decay using SGD, Adam, and K-FAC across several neural-network architectures. They concluded that weight decay does not operate through one universal mechanism. Instead, its effect depends on both the optimizer and the architecture [@zhang2018]. One of the mechanisms is increasing the effective learning rate.

For layers followed by Batch Normalization, multiplying the weights by a positive constant does not change the represented network function:

$$
\mathrm{BN}(x;\eta\boldsymbol{\theta}_l)
=
\mathrm{BN}(x;\boldsymbol{\theta}_l),
\qquad \eta>0.
$$

Consequently,  rescaling the norm of these weights does not meaningfully constrain the reprsentation. Nevertheless, weight decay can still improve generalization.

For a scale-invariant weight vector, the approximate effective learning rate for its direction is proportional to

$$
\frac{\alpha}
{\left\|
\boldsymbol{\theta}
\right\|_2^2}.
$$

By reducing the weight norm, weight decay increases the effective learning rate. This may strengthen the regularizing effect of stochastic gradient noise.

Zhang et al. found that, for Batch-Normalized networks trained with SGD or Adam, most of the generalization improvement came from applying weight decay to the normalized convolutional layers rather than only to the final unnormalized layer. They also reproduced almost the full effect of weight decay by matching the weight norms, and therefore the effective learning-rate schedule, in a network trained without weight decay [@zhang2018].

### Source Map

| Claim | Exact primary source and locator | What it establishes | What it does not establish |
|---|---|---|---|
| Under standard SGD, an $L_2$ penalty and direct weight decay generate identical updates after rescaling the coefficients by the learning rate. | [@loshchilov2019], Section 2, Equation (1) and Proposition 1. | It algebraically establishes equivalence when $\lambda_{WD}=\alpha\lambda_{L2}$. | It does not establish equivalence when the coefficients are not rescaled or when the optimizer uses nonuniform adaptive preconditioning. |
| For adaptive gradient methods, an ordinary $L_2$ penalty is generally not equivalent to direct weight decay. | [@loshchilov2019], Section 2, Algorithm 2 and Proposition 2. | It shows that adaptive optimization also transforms the $L_2$ penalty gradient, whereas decoupled weight decay shrinks parameters separately from the adaptive gradient update. | It does not imply that AdamW must outperform Adam for every task, dataset, architecture, or hyperparameter configuration. |
| AdamW improved generalization and made the learning-rate and weight-decay hyperparameters more separable in the reported experiments. | [@loshchilov2019], Section 4.1--4.3. | It provides empirical evidence on ResNets trained on CIFAR-10 and ImageNet32x32 under several learning-rate schedules and training budgets. | It does not establish universal superiority beyond the tested image-classification settings. |
| The regularization effect of weight decay depends on the optimizer and network architecture rather than following one universal mechanism. | [@zhang2018], Abstract and Introduction.| The study identifies three mechanisms: increasing effective learning rate, approximately reducing the input-output Jacobian norm, and reducing effective damping in second-order optimization. | It does not establish that these are the only possible mechanisms or that each mechanism applies to every neural network. |
| In Batch-Normalized networks trained with SGD or Adam, weight decay acts primarily by increasing the effective learning rate. | [@zhang2018], Section 4.1. | Because Batch-Normalized layers are scale-invariant, decreasing the weight norm raises the effective learning rate. Matching the weight norms reproduced almost the entire observed generalization effect. | It does not show that the same explanation applies to networks without normalization or to optimizers whose updates are invariant to parameter rescaling. |
| Dropout is a stochastic regularization method that temporarily removes randomly selected units and their incoming and outgoing connections during training. | [@srivastava2014dropout], Section 1. | The paper defines dropout as randomly and temporarily dropping units and their connections during training. Each training case therefore uses a sampled "thinned" network. | The source does not establish that the dropped units or connections are permanently removed from the trained model. | 
| Magnitude-based network pruning permanently removes connections whose weights fall below a selected threshold and then retrains the remaining sparse connections. | [@han2016deep], Section 2. | The paper describes a three-step pruning procedure: train the original network, remove small-weight connections below a threshold, and retrain the remaining sparse network. | The source does not establish that every pruning rule, threshold, architecture, or data set will preserve predictive accuracy. |
| A multilayer neural network can use backpropagation to adjust connection weights in the output and successively earlier hidden layers, enabling hidden units to learn representations relevant to the task. | [@Rumelhart1986LearningRB], Abstract; p. 534, description of the backward pass; p. 535, Eq. (8) and the paragraph immediately preceding it. | The paper establishes that errors can be propagated backward through a multilayer network to adjust the weights and learn internal representations. | It does not establish which regularization method, such as Ridge, Lasso, or Elastic Net, performs best in all deep-network architectures or tasks. |


## 4B. Elastic net on original predictors

```{r original-elastic-net, eval=FALSE}
alpha_grid <- c(0.10, 0.25, 0.50, 0.75, 0.90)

# Fit one cv.glmnet object per alpha with the same foldid.
# Select alpha and lambda using training CV only.

fit_elastic_net_4B <- function(x, y, foldid, alpha_list){
  fits <- list()
  for (alpha in alpha_list){
        fits[[as.character((alpha))]] <- cv.glmnet(
        x = x,
        y = y,
        alpha = alpha,
        foldid = foldid
    )
  }
  return(fits)
}

plot_coef_path_4B <- function(glmnet_fit, alpha, title) {

  if (!is.null(alpha)) {
    title <- paste0(title, " (alpha = ", alpha, ")")
  } 
  
  coef_mat <- as.matrix(glmnet_fit$beta)
  colnames(coef_mat) <- glmnet_fit$lambda

  coef_df <- as.data.frame(coef_mat) |>
    tibble::rownames_to_column("variable") |>
    tidyr::pivot_longer(
      -variable,
      names_to = "lambda",
      values_to = "coefficient"
    ) |>
    dplyr::mutate(
      lambda = as.numeric(lambda)
    )

  ggplot2::ggplot(
    coef_df,
    ggplot2::aes(
      x = lambda,
      y = coefficient,
      colour = variable
    )
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = title,
      x = expression(lambda),
      y = "Coefficient"
    ) +
    ggplot2::theme_minimal()
}

plot_nonzero_slopes_4B <- function(glmnet_fit, alpha, title) {
  
  if (!is.null(alpha)) {
    title <- paste0(title, " (alpha = ", alpha, ")")
  } 

  df_data <- tibble::tibble(
    lambda = glmnet_fit$lambda,
    df = glmnet_fit$df
  )

  ggplot2::ggplot(
    df_data,
    ggplot2::aes(
      x = lambda,
      y = df,
    )
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = title,
      x = expression(lambda),
      y = "number of none-zero slopes"
    ) +
    ggplot2::theme_minimal()
}

plot_cross_validation_error <- function(cv_glmnet_fit, alpha, title) {
  
  if (!is.null(alpha)) {
    title <- paste0(title, " (alpha = ", alpha, ")")
  } 

  error_data <- tibble::tibble(
    lambda = cv_glmnet_fit$lambda,
    error = cv_glmnet_fit$cvm
  )

  ggplot2::ggplot(
    error_data,
    ggplot2::aes(
      x = lambda,
      y = error,
    )
  ) +
    ggplot2::geom_line() +
    ggplot2::labs(
      title = title,
      x = expression(lambda),
      y = "cross-validation error"
    ) +
    ggplot2::theme_minimal()
}

elastic_net_list <- fit_elastic_net_4B(x_train, y_train, foldid, alpha_grid)

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]$glmnet.fit
  print(plot_coef_path_4B(fit, alpha, "Elastic Net Coefficient Paths"))
}

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]$glmnet.fit
  print(plot_nonzero_slopes_4B(fit, alpha, "Number of Non-Zero Slopes"))
}

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]
  print(plot_cross_validation_error(fit, alpha, "Cross-Validation Error"))
}

cv_results <- tibble::tibble(
  alpha = numeric(),
  lambda = numeric(),
  cv_error = numeric()
)

for (alpha in names(elastic_net_list)) {
  fit <- elastic_net_list[[alpha]]
  idx <- which(fit$lambda == fit$lambda.min)
  error <- fit$cvm[idx]
  cv_results <- dplyr::add_row(
    cv_results,
    alpha = as.numeric(alpha),
    lambda = fit$lambda.min,
    cv_error = error
  )
}

best_elastic_net_param <- cv_results |> dplyr::slice_min(cv_error, n=1)
best_elastic_net_param <- best_elastic_net_param[1, ]
best_elastic_net_model <- elastic_net_list[[as.character(
  best_elastic_net_param$alpha
)]]
```



## 4C. Fixed ReLU features

```{r random-relu-features, eval=FALSE}
relu <- function(z, A, C){
  z <- z %*% t(A)
  C <- matrix(C, nrow = nrow(z), ncol = length(C), byrow = TRUE)
  z <- z + C
  return (pmax(z, 0))
} 
M <- 200L
p <- ncol(x_train)

# TODO:
# 1. Generate A and bias once from seeds$features using the stated distributions.
# 2. Apply the same A and bias to x_train and x_test.
# 3. Fit a scaler to H_train only and apply it to H_test.
# 4. Fit ridge, lasso, and elastic net on H_train with the shared foldid.

set.seed(22)

A <- matrix(rnorm(M * p, mean = 0, sd = sqrt(1/p)),
            nrow = M, ncol = p)

C <- matrix(rnorm(M * 1, mean = 0, sd = 0.5 ),
            nrow = M, ncol = 1)

x_hidden <- relu(x_train, A, C)
same_col <- apply(x_hidden, 2, function(col) all(col == col[1]))
print(same_col)

cv_ridge <- cv.glmnet(
  x = x_hidden,
  y = y_train,
  alpha = 0,     
  foldid = foldid 
)

cv_lasso <- cv.glmnet(
  x = x_hidden,
  y = y_train,
  alpha = 1,     
  foldid = foldid 
)

cv_elastic_net_list <- fit_elastic_net_4B(x_hidden, y_train, foldid, alpha_grid)

plot_cross_validation_error(cv_ridge, NULL, "Cross Validation Error of Ridge")
plot_nonzero_slopes_4B(cv_ridge$glmnet.fit, NULL, "Number of Active Hidden Features of Ridge")

plot_cross_validation_error(cv_lasso, NULL, "Cross Validation Error of Lasso")
plot_nonzero_slopes_4B(cv_lasso$glmnet.fit, NULL, "Number of Active Hidden Features of Lasso")

for (alpha in names(cv_elastic_net_list)) {
  fit <- cv_elastic_net_list[[alpha]]$glmnet.fit
  print(plot_nonzero_slopes_4B(fit, alpha, "Number of Non-Zero Slopes of Elastic Net"))
}

for (alpha in names(cv_elastic_net_list)) {
  fit <- cv_elastic_net_list[[alpha]]
  print(plot_cross_validation_error(fit, alpha, "Cross-Validation Error of Elastic Net"))
}

cv_results <- tibble::tibble(
  alpha = numeric(),
  lambda = numeric(),
  cv_error = numeric()
)

for (alpha in names(cv_elastic_net_list)) {
  fit <- cv_elastic_net_list[[alpha]]
  idx <- which(fit$lambda == fit$lambda.min)
  error <- fit$cvm[idx]
  cv_results <- dplyr::add_row(
    cv_results,
    alpha = as.numeric(alpha),
    lambda = fit$lambda.min,
    cv_error = error
  )
}

best_elastic_net_param_4C <- cv_results |> dplyr::slice_min(cv_error, n=1)
best_elastic_net_param_4C <- best_elastic_net_param_4C[1, ]
best_elastic_net_model_4C <- cv_elastic_net_list[[as.character(
  best_elastic_net_param_4C$alpha
)]]$glmnet.fit

best_ridge_model_4C <- cv_ridge
best_lasso_model_4C <- cv_lasso

```

*State which weights are fixed, which are learned, and how active features are counted.*

## 4D. Locked declaration and final holdout evaluation

> **Locked before viewing `y_test`:** [preprocessing, candidate specifications, selected tuning values, feature seed, nominated deployment model, metrics]

```{r final-holdout-evaluation, eval=FALSE}
# This must be the first analysis chunk that uses y_test for evaluation.
# Generate all fixed candidate predictions and one RMSE/MAE table here.

MAE <- function(actual, predicted) {
  mean(abs(actual - predicted))
}

RMSE <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}


```

*State whether the holdout evidence supports the predeclared choice. Do not retune.*

## 4E. Deep-learning connection and limitations

The fixed-feature experiment can be written as

\[
\hat{y}(x)
=
\beta_0
+
\sum_{m=1}^{M} b_m h_m(x),
\qquad
h_m(x)
=
\max\{a_m^\top x+c_m,0\}.
\]

The hidden-feature parameters \(a_m\) and \(c_m\) were generated once and then held fixed. Therefore, only the intercept \(\beta_0\) and the output weights \(b_1,\ldots,b_M\) were estimated from the training data.

### \(L_2\) regularization and weight decay

Applying an \(L_2\) penalty to the learned output weights gives the objective

\[
\frac{1}{2n}\lVert y-Hb\rVert_2^2
+
\frac{\lambda}{2}\lVert b\rVert_2^2,
\]

where \(H\) is the fixed ReLU feature matrix. This is exactly a Ridge-regression objective on the hidden features because the squared output weights are added directly to the squared-error loss.

For adaptive optimizers, the gradient of the \(L_2\) term is scaled by the optimizer's parameter-specific adaptive factors. Direct weight decay instead shrinks the parameters separately from the adaptive loss-gradient update [@loshchilov2019]. Thus, the \(L_2\) penalty used in this experiment is a Ridge penalty on the output weights, but in adaptive neural-network optimization, it is not always identical to weight decay.

### Zero output weights and feature sparsity

If a fitted output weight satisfies

\[
b_m=0,
\]

then the corresponding term \(b_mh_m(x)\) contributes zero to every prediction. In that sense, the \(m\)-th fixed hidden feature is removed from this scalar predictor.

Nevertheless, this does not create sparsity in the input-to-hidden parameters. The vector \(a_m\), which defines how the original predictors are combined before the ReLU transformation, remains unchanged and may still contain nonzero entries for every original predictor. Therefore, Lasso or Elastic net can produce sparsity among the learned output weights, but this experiment does not produce sparse input-to-hidden connections.

### Dropout and permanent pruning

Dropout temporarily applies random masks to hidden activations during training. A unit may be inactive for one training update and active again in a later update. The purpose is to train across many randomly thinned network configurations rather than to delete a fixed set of units permanently [@srivastava2014dropout].

Permanent pruning is different because selected weights, connections, or units are removed from the model and remain absent during subsequent inference or retraining [@han2016deep]. A zero output weight in the present experiment is closer to permanent removal of one fixed feature than to dropout, although it still affects only the connection from that feature to the scalar output.

### Scope of the neural-network connection

This experiment contains only one random hidden layer. The input-to-hidden weights and hidden biases were sampled once and frozen, while only the linear output layer was trained. It is therefore better interpreted as regularized regression on a fixed nonlinear representation than as a fully trained neural network.

A multilayer neural network can use backpropagation to adjust connection weights in the output and successively earlier hidden layers, enabling hidden units to learn representations relevant to the task [@Rumelhart1986LearningRB]. Consequently, the present results do not establish which regularization method would perform best in a fully trained deep network.


### Limitations

First, the results are sensitive to the random-feature seed. A different seed would generate different vectors \(a_m\) and offsets \(c_m\), producing a different hidden-feature matrix. This could change the number of constant features, the selected tuning parameters, the cross-validation errors, and the number of active hidden features. Because only one declared seed was used, the experiment does not measure stability across alternative random representations.

Second, the hidden representation was not learned from the response. Random ReLU features may capture useful nonlinear structure, but they are not optimized for this specific prediction task. A fully trained network could adapt its hidden weights and biases to construct a more task-relevant representation.

Third, the choice \(M=200\) was fixed rather than tuned. A small value of \(M\) might provide insufficient nonlinear flexibility, whereas a large value could increase computational cost and model variance. Therefore, the conclusions may depend on the selected number of hidden features.


# Problem 5. Report quality and reproducibility

## 5A. Reproduction record

```text
1. Open a clean R session.
2. [Exact commands and folder layout.]
3. Render GroupXX_ProjectQuiz1.Rmd.
```

## 5B. Recommendation and limitations

*Give the final recommendation, distinguish prediction from interpretation, and state limitations.*

## 5C. AI verification record

| AI-assisted item | How it was checked | Evidence | Modification or rejection |
|---|---|---|---|
| Drafting the source map for dropout and network pruning | Each claim was checked against the original research papers and the cited section, page, or figure. | Srivastava et al. (2014), Section 1; Han et al. (2016), Section 2. | The wording was revised to avoid claims beyond the sources. |
| Explaining the difference between $L_2$ regularization and weight decay | The explanation and equations were compared with the propositions and algorithms in the original paper. | Loshchilov and Hutter (2019), Section 2. | Overgeneralized statements were rejected. |
# Preflight and session information {-}

```{r preflight, eval=FALSE}
# Add every object required to reproduce the final table and figures.
required_objects <- c(
  "train_id", "test_id", "x_train", "x_test", "y_train", "y_test", "foldid"
)
missing_objects <- setdiff(required_objects, ls())
if (length(missing_objects)) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}
stopifnot(length(intersect(train_id, test_id)) == 0L)
```

```{r session-info}
sessionInfo()
```
