---
title: "ASESII 24A02 Project-Based Quiz 1"
subtitle: "Ridge, Lasso, and Regularization for Neural Features"
author: "Group 02: Nguyễn Thanh An - 24125024, Nguyễn Chí Bảo - 24125025, Trần Nguyễn Khánh Bình - 24125026, Phạm Văn Trang - 24125046, Quách Thiên Lạc - 24125092"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    fig_crop: false
  html_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
link-citations: true
params:
  group_number: 2
fontsize: 11pt
geometry: margin=1in
---

<!--
Copy this template into your group submission folder, rename it
Group02_ProjectQuiz1.Rmd, place references.bib beside it, and set the group
number parameter before rendering. Remove instructional comments from the
final report.
-->

# Authorship and contribution statement {-}

We confirm that every group member can explain the submitted analysis. AI use is reported in `Group02_AI_Log.csv`, and the group checked every AI-assisted item used in this report.

| Student | ID | Main contributions | Verification performed |
|---|---|---|---|
| Nguyễn Thanh An        | 24125024 |---|---|
| Nguyễn Chí Bảo         | 24125025 |---|---|
| Trần Nguyễn Khánh Bình | 24125026 |---|---|
| Phạm Văn Trang         | 24125046 |---|---|
| Quách Thiên Lạc        | 24125092 |---|---|


# Abstract {-}

*Maximum 150 words: prediction problem, methods, main evidence, and recommendation.*

# Reproducible setup {-}

```{r setup, message=FALSE, warning=FALSE}
required_packages <- c("tidyverse", "glmnet", "broom", "knitr")
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop("Install the required packages before rendering: ",
       paste(missing_packages, collapse = ", "))
}

library(tidyverse)
library(glmnet)
library(broom)
library(knitr)

# hardcode group number bc somehow the object group_number in param is not found
group_number <- 2

# group_number <- as.integer(params$group_number)
# if (length(group_number) != 1L || is.na(group_number) || group_number < 1L) {
#   stop("Set params$group_number to your assigned positive group number.")
# }

seeds <- list(
  split = 240200L + group_number,
  cv = 240300L + group_number,
  features = 240400L + group_number
)
```

## Shared helper functions {-}

```{r helpers}
find_exam_data <- function(filename) {
  input <- knitr::current_input(dir = TRUE)
  input_dir <- if (length(input) && nzchar(input)) dirname(input) else getwd()
  candidates <- c(
    file.path(input_dir, "data", filename),
    file.path(input_dir, "..", "data", filename),
    file.path(getwd(), "data", filename)
  )
  existing <- candidates[file.exists(candidates)]
  if (!length(existing)) stop("Cannot locate ", filename, " by a relative path.")
  hits <- unique(normalizePath(existing, winslash = "/", mustWork = TRUE))
  if (length(hits) > 1L && length(unique(unname(tools::md5sum(hits)))) > 1L) {
    stop("Multiple nonidentical copies of ", filename, " were found.")
  }
  hits[[1L]]
}

split_rows <- function(n, train_prop = 0.80, seed) {
  stopifnot(n >= 10L, train_prop > 0, train_prop < 1)
  set.seed(seed)
  train <- sort(sample.int(n, floor(train_prop * n), replace = FALSE))
  list(train = train, test = setdiff(seq_len(n), train))
}

fit_scaler <- function(x, tol = 1e-12) {
  x <- as.matrix(x)
  center <- colMeans(x)
  centered <- sweep(x, 2L, center, "-")
  scale <- sqrt(colMeans(centered^2))
  keep <- is.finite(scale) & scale > tol
  if (!any(keep)) stop("No nonconstant training predictor remains.")
  list(
    center = center[keep],
    scale = scale[keep],
    keep = keep,
    dropped = colnames(x)[!keep]
  )
}

apply_scaler <- function(x, prep) {
  x <- as.matrix(x)[, prep$keep, drop = FALSE]
  x <- sweep(x, 2L, prep$center, "-")
  sweep(x, 2L, prep$scale, "/")
}

make_foldid <- function(n, k = 5L, seed) {
  if (k < 3L || k > n) stop("Require 3 <= k <= number of training rows.")
  set.seed(seed)
  sample(rep(seq_len(k), length.out = n))
}

fit_cv_glmnet <- function(x, y, alpha, foldid) {
  glmnet::cv.glmnet(
    x = x,
    y = y,
    family = "gaussian",
    alpha = alpha,
    foldid = foldid,
    standardize = FALSE,
    intercept = TRUE,
    type.measure = "mse"
  )
}

score_regression <- function(y, prediction) {
  y <- as.numeric(y)
  prediction <- as.numeric(prediction)
  if (length(y) != length(prediction) || any(!is.finite(c(y, prediction)))) {
    stop("Metrics require equal-length finite vectors.")
  }
  c(RMSE = sqrt(mean((y - prediction)^2)),
    MAE = mean(abs(y - prediction)))
}

count_nonzero <- function(fit, s = "lambda.min", tol = 1e-8) {
  b <- as.matrix(stats::coef(fit, s = s))
  slopes <- b[rownames(b) != "(Intercept)", 1L]
  sum(abs(slopes) > tol)
}

safe_condition_numbers <- function(x, lambda, tol = 1e-10) {
  eigenvalues <- eigen(
    crossprod(x) / nrow(x),
    symmetric = TRUE,
    only.values = TRUE
  )$values
  largest <- max(eigenvalues)
  smallest <- max(min(eigenvalues), 0)
  cutoff <- tol * max(1, largest)
  c(
    gram = if (smallest <= cutoff) Inf else largest / smallest,
    ridge = (largest + lambda) / (smallest + lambda)
  )
}
```

# Problem 1. Prediction design and leakage control

## 1A. Target and predictors

```{r data-configuration}
config <- if (group_number %% 2L == 1L) {
  list(
    file = "fat.csv",
    response = "brozek",
    excluded = c("brozek", "siri", "density", "free")
  )
} else {
  list(
    file = "prostate.csv",
    response = "lpsa",
    excluded = "lpsa"
  )
}

data_raw <- read.csv(find_exam_data(config$file), check.names = FALSE)
predictors <- setdiff(names(data_raw), config$excluded)
analysis_data <- data_raw[, c(config$response, predictors), drop = FALSE]

if (!all(vapply(analysis_data, is.numeric, logical(1)))) {
  stop("The assigned response and predictors must be numeric.")
}

# csv output for preflight check
cat("File:", config$file, "\n")
cat("Response:", config$response, "\n")
cat("Rows:", nrow(data_raw), "\n")
cat("Columns:", ncol(data_raw), "\n")

# basic checks 
integrity_summary <- tibble::tibble(
    variable = names(data_raw),
    type = vapply(
        data_raw,
        function(x) class(x)[1],
        character(1)
    ),
    missing = vapply(
        data_raw,
        function(x) sum(is.na(x)),
        integer(1)
    ),
    unique_values = vapply(
        data_raw,
        function(x) dplyr::n_distinct(
            x,
            na.rm = TRUE
        ),
        integer(1)
    )
)

knitr::kable(
    integrity_summary,
    caption = "Basic CSV integrity checks"
)
```

*Define the prediction task and justify every exclusion.*

## 1B. Split and train-only preprocessing

```{r split-and-scale}
# Split once using the declared group-specific seed.
split <- split_rows(
    n = nrow(analysis_data),
    train_prop = 0.80,
    seed = seeds$split
)

train_id <- split$train
test_id <- split$test

stopifnot(
    length(intersect(train_id, test_id)) == 0L
)

# Preserve unprocessed train/test objects for training-only exploration.
x_train_raw <- analysis_data[
    train_id,
    predictors,
    drop = FALSE
]

x_test_raw <- analysis_data[
    test_id,
    predictors,
    drop = FALSE
]

y_train_raw <- analysis_data[
    train_id,
    config$response,
    drop = TRUE
]

# Keep the supplied holdout response locked on its original lpsa scale.
y_test_locked <- analysis_data[
    test_id,
    config$response,
    drop = TRUE
]

# Apply the fixed transformation map to predictors. (only change age as per newest decision)
transform_predictors <- function(x) {
    transformed <- x
    transformed$age <- x$age^2
    transformed$lbph <- transformed$lpbh
    transformed$lcavol <- x$lcavol
    transformed$lcp <- transformed$lcp
    transformed$lweight <- x$lweight
    transformed$pgg45 <- transformed$pgg45
    transformed$gleason <- x$gleason
    transformed$svi <- x$svi
    transformed
}

x_train_transformed <- transform_predictors(x_train_raw)
x_test_transformed <- transform_predictors(x_test_raw)

# Transform the training response from log(PSA) to sqrt(PSA).
# Reuse this function on y_test_locked only in the final evaluation chunk.
transform_response <- function(y) {
    exp(y / 2)
}

y_train <- transform_response(y_train_raw)

# Fit preprocessing constants on transformed training predictors only.
x_prep <- fit_scaler(
    x_train_transformed
)

x_train <- apply_scaler(
    x_train_transformed,
    x_prep
)

x_test <- apply_scaler(
    x_test_transformed,
    x_prep
)

# Create one shared five-fold assignment for all training comparisons.
foldid <- make_foldid(
    n = nrow(x_train),
    k = 5L,
    seed = seeds$cv
)

cat("\nFold sizes:\n")
print(table(foldid))

# Record the fixed rules used by the pipeline.
transformation_record <- tibble::tibble(
    variable = c(predictors, config$response),
    role = c(rep("predictor", length(predictors)), "response")
) |>
    dplyr::mutate(
        transformation = dplyr::case_when(
            variable == "age" ~ "x^2",
            variable == "lbph" ~ "none",
            variable == "lcavol" ~ "none",
            variable == "lcp" ~ "none",
            variable == "lweight" ~ "none",
            variable == "pgg45" ~ "none",
            variable == "gleason" ~ "none",
            variable == "svi" ~ "none",
            variable == "lpsa" ~ "exp(x / 2) = sqrt(PSA)",
            TRUE ~ "none"
        )
    )

knitr::kable(
    transformation_record,
    caption = "Fixed transformation rules"
)
```

*Report seeds, dimensions, scaling, folds, and leakage safeguards.*

## 1C. Training-data exploration

All exploratory calculations in this section use training observations only. The holdout response remains locked.

```{r training-eda, fig.width=9, fig.height=6}
# ---------------------------------------------------------
# Output directories
# ---------------------------------------------------------

plot_directory <- file.path(
    "submission/submission_data",
    "plots"
)

table_directory <- file.path(
    "submission/submission_data",
    "tables"
)

dir.create(
    plot_directory,
    recursive = TRUE,
    showWarnings = FALSE
)

dir.create(
    table_directory,
    recursive = TRUE,
    showWarnings = FALSE
)


# ---------------------------------------------------------
# Helper function
# ---------------------------------------------------------

sample_skewness <- function(x) {
    x <- x[is.finite(x)]
    n <- length(x)

    if (n < 3L) {
        return(NA_real_)
    }

    standard_deviation <- stats::sd(x)

    if (
        !is.finite(standard_deviation) ||
        standard_deviation == 0
    ) {
        return(NA_real_)
    }

    standardized <- (
        x - mean(x)
    ) / standard_deviation

    n / (
        (n - 1) * (n - 2)
    ) * sum(standardized^3)
}


# ---------------------------------------------------------
# Variable groups
# ---------------------------------------------------------

continuous_predictors <- predictors[
    vapply(
        x_train_raw,
        function(x) {
            is.numeric(x) &&
                dplyr::n_distinct(
                    x,
                    na.rm = TRUE
                ) > 5L
        },
        logical(1)
    )
]

discrete_predictors <- intersect(
    c(
        "svi",
        "gleason"
    ),
    predictors
)

# Age is shown separately in the transformation comparison.
continuous_predictors_as_is <- setdiff(
    continuous_predictors,
    "age"
)


# ---------------------------------------------------------
# 1. Response before and after transformation
# ---------------------------------------------------------

response_plot_data <- dplyr::bind_rows(
    tibble::tibble(
        version = "Before: lpsa",
        value = y_train_raw
    ),
    tibble::tibble(
        version = "After: sqrt(PSA)",
        value = y_train
    )
) |>
    dplyr::mutate(
        version = factor(
            version,
            levels = c(
                "Before: lpsa",
                "After: sqrt(PSA)"
            )
        )
    )

response_transformation_plot <- ggplot2::ggplot(
    response_plot_data,
    ggplot2::aes(x = value)
) +
    ggplot2::geom_histogram(
        bins = 20,
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::facet_wrap(
        ~ version,
        scales = "free",
        ncol = 2
    ) +
    ggplot2::labs(
        title = "Response before and after transformation",
        subtitle = paste(
            "Training observations only;",
            "sqrt(PSA) = exp(lpsa / 2)"
        ),
        x = NULL,
        y = "Count"
    ) +
    ggplot2::theme_minimal()

print(response_transformation_plot)

ggplot2::ggsave(
    filename = file.path(
        plot_directory,
        "1. response_transformation_comparison.png"
    ),
    plot = response_transformation_plot,
    width = 9,
    height = 5,
    units = "in",
    dpi = 300,
    bg = "white"
)


# ---------------------------------------------------------
# 2. Age before and after transformation
# ---------------------------------------------------------

age_plot_data <- dplyr::bind_rows(
    tibble::tibble(
        version = "Before: age",
        value = x_train_raw$age
    ),
    tibble::tibble(
        version = "After: age^2",
        value = x_train_transformed$age
    )
) |>
    dplyr::mutate(
        version = factor(
            version,
            levels = c(
                "Before: age",
                "After: age^2"
            )
        )
    )

age_transformation_plot <- ggplot2::ggplot(
    age_plot_data,
    ggplot2::aes(x = value)
) +
    ggplot2::geom_histogram(
        bins = 20,
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::facet_wrap(
        ~ version,
        scales = "free_x",
        ncol = 2
    ) +
    ggplot2::labs(
        title = "Age before and after transformation",
        subtitle = "Training observations only",
        x = NULL,
        y = "Count"
    ) +
    ggplot2::theme_minimal()

print(age_transformation_plot)

ggplot2::ggsave(
    filename = file.path(
        plot_directory,
        "2. age_transformation_comparison.png"
    ),
    plot = age_transformation_plot,
    width = 9,
    height = 5,
    units = "in",
    dpi = 300,
    bg = "white"
)


# ---------------------------------------------------------
# 3. Skewness comparison for age and response only
# ---------------------------------------------------------

transformation_skewness_table <- tibble::tibble(
    variable = c(
        "age",
        "response"
    ),
    transformation = c(
        "age^2",
        "exp(lpsa / 2) = sqrt(PSA)"
    ),
    original_skewness = c(
        sample_skewness(
            x_train_raw$age
        ),
        sample_skewness(
            y_train_raw
        )
    ),
    transformed_skewness = c(
        sample_skewness(
            x_train_transformed$age
        ),
        sample_skewness(
            y_train
        )
    )
) |>
    dplyr::mutate(
        original_absolute_skewness =
            abs(original_skewness),
        transformed_absolute_skewness =
            abs(transformed_skewness),
        absolute_skewness_reduction =
            original_absolute_skewness -
            transformed_absolute_skewness,
        result = dplyr::if_else(
            absolute_skewness_reduction > 0,
            "Reduced absolute skewness",
            "Did not reduce absolute skewness"
        )
    )

knitr::kable(
    transformation_skewness_table,
    digits = 3,
    caption = "Age and response skewness before and after transformation"
)

readr::write_csv(
    transformation_skewness_table,
    file = file.path(
        table_directory,
        "3. transformation_skewness_comparison.csv"
    )
)


# ---------------------------------------------------------
# 4. Remaining continuous predictors on supplied scales
# ---------------------------------------------------------

original_continuous_plot_data <- x_train_raw |>
    dplyr::select(
        dplyr::all_of(
            continuous_predictors_as_is
        )
    ) |>
    tidyr::pivot_longer(
        cols = dplyr::everything(),
        names_to = "variable",
        values_to = "value"
    )

original_continuous_predictor_plot <- ggplot2::ggplot(
    original_continuous_plot_data,
    ggplot2::aes(x = value)
) +
    ggplot2::geom_histogram(
        bins = 20,
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::facet_wrap(
        ~ variable,
        scales = "free_x",
        ncol = 3
    ) +
    ggplot2::labs(
        title = "Original continuous predictor distributions",
        subtitle = paste(
            "Training observations only;",
            "age is shown separately"
        ),
        x = NULL,
        y = "Count"
    ) +
    ggplot2::theme_minimal()

print(original_continuous_predictor_plot)

ggplot2::ggsave(
    filename = file.path(
        plot_directory,
        "4. original_continuous_predictor_distributions.png"
    ),
    plot = original_continuous_predictor_plot,
    width = 9,
    height = 6,
    units = "in",
    dpi = 300,
    bg = "white"
)


# ---------------------------------------------------------
# 5. Discrete predictors on supplied scales
# ---------------------------------------------------------

discrete_plot_data <- x_train_raw |>
    dplyr::select(
        dplyr::all_of(
            discrete_predictors
        )
    ) |>
    tidyr::pivot_longer(
        cols = dplyr::everything(),
        names_to = "variable",
        values_to = "value"
    ) |>
    dplyr::count(
        variable,
        value,
        name = "count"
    )

original_discrete_predictor_plot <- ggplot2::ggplot(
    discrete_plot_data,
    ggplot2::aes(
        x = factor(value),
        y = count
    )
) +
    ggplot2::geom_col(
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::facet_wrap(
        ~ variable,
        scales = "free_x"
    ) +
    ggplot2::labs(
        title = "Original discrete predictor distributions",
        subtitle = "Training observations only",
        x = "Predictor value",
        y = "Count"
    ) +
    ggplot2::theme_minimal()

print(original_discrete_predictor_plot)

ggplot2::ggsave(
    filename = file.path(
        plot_directory,
        "5. original_discrete_predictor_distributions.png"
    ),
    plot = original_discrete_predictor_plot,
    width = 7,
    height = 5,
    units = "in",
    dpi = 300,
    bg = "white"
)


# ---------------------------------------------------------
# 6. Training correlation table
# ---------------------------------------------------------

correlation_data <- x_train_transformed |>
    dplyr::mutate(
        response = y_train,
        .before = 1
    )

correlation_matrix <- stats::cor(
    correlation_data,
    use = "complete.obs"
)

knitr::kable(
    round(correlation_matrix, 3),
    caption = paste(
        "Training correlations after transformation",
        "and before scaling"
    )
)

readr::write_csv(
    tibble::as_tibble(
        correlation_matrix,
        rownames = "variable"
    ),
    file = file.path(
        table_directory,
        "6. training_correlation_matrix.csv"
    )
)


# ---------------------------------------------------------
# 7. Predictor-response relationship
# ---------------------------------------------------------
# Extract correlations between every predictor and the response.

predictor_response_plot_data <- x_train_transformed |>
    dplyr::mutate(
        response = y_train
    ) |>
    tidyr::pivot_longer(
        cols = -response,
        names_to = "predictor",
        values_to = "predictor_value"
    )


# Calculate the correlation shown in each facet label.

predictor_response_correlations <-
    predictor_response_plot_data |>
    dplyr::group_by(
        predictor
    ) |>
    dplyr::summarise(
        correlation = stats::cor(
            predictor_value,
            response,
            use = "complete.obs"
        ),
        .groups = "drop"
    )

correlation_labels <- stats::setNames(
    paste0(
        predictor_response_correlations$predictor,
        " (r = ",
        sprintf(
            "%.3f",
            predictor_response_correlations$correlation
        ),
        ")"
    ),
    predictor_response_correlations$predictor
)


# Create the faceted scatterplot.

all_predictor_response_plot <- ggplot2::ggplot(
    predictor_response_plot_data,
    ggplot2::aes(
        x = predictor_value,
        y = response
    )
) +
    ggplot2::geom_point(
        color = "steelblue",
        alpha = 0.7,
        size = 1.8
    ) +
    ggplot2::geom_smooth(
        method = "lm",
        formula = y ~ x,
        se = TRUE,
        color = "darkred",
        fill = "grey75",
        linewidth = 0.9
    ) +
    ggplot2::facet_wrap(
        ~ predictor,
        scales = "free_x",
        ncol = 2,
        labeller = ggplot2::as_labeller(
            correlation_labels
        )
    ) +
    ggplot2::labs(
        title = "Predictors and transformed response",
        subtitle = paste(
            "Training observations only;",
            "red lines are fitted linear relationships"
        ),
        x = "Predictor value",
        y = "sqrt(PSA)"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
        strip.text = ggplot2::element_text(
            face = "bold"
        )
    )

print(
    all_predictor_response_plot
)


# Save the plot.

ggplot2::ggsave(
    filename = file.path(
        plot_directory,
        "7. all_predictors_vs_sqrt_psa.png"
    ),
    plot = all_predictor_response_plot,
    width = 10,
    height = 12,
    units = "in",
    dpi = 300,
    bg = "white"
)


# Export the correlations used in the facet labels.

readr::write_csv(
    predictor_response_correlations,
    file = file.path(
        table_directory,
        "predictor_response_correlations.csv"
    )
)

# ---------------------------------------------------------
# 7.5. Pairwise feature covariance matrix
# ---------------------------------------------------------

feature_covariance_matrix <- stats::cov(
    x_train_transformed,
    use = "complete.obs"
)


# Convert the matrix to a table for display and CSV export.

feature_covariance_table <- feature_covariance_matrix |>
    as.data.frame() |>
    tibble::rownames_to_column(
        var = "predictor"
    )

knitr::kable(
    feature_covariance_table,
    digits = 3,
    caption = paste(
        "Pairwise covariance matrix of transformed",
        "training predictors"
    )
)


# Export the covariance matrix to CSV.

readr::write_csv(
    feature_covariance_table,
    file = file.path(
        table_directory,
        "feature_covariance_matrix.csv"
    )
)


# Convert the covariance matrix to long format for PNG output.

covariance_image_data <- as.data.frame(
    as.table(
        feature_covariance_matrix
    )
) |>
    tibble::as_tibble() |>
    dplyr::rename(
        row_predictor = Var1,
        column_predictor = Var2,
        covariance = Freq
    ) |>
    dplyr::mutate(
        row_predictor = factor(
            row_predictor,
            levels = rev(
                rownames(feature_covariance_matrix)
            )
        ),
        column_predictor = factor(
            column_predictor,
            levels = colnames(
                feature_covariance_matrix
            )
        ),
        displayed_value = formatC(
            covariance,
            format = "g",
            digits = 3
        )
    )


# Create a table-like image.

covariance_table_plot <- ggplot2::ggplot(
    covariance_image_data,
    ggplot2::aes(
        x = column_predictor,
        y = row_predictor
    )
) +
    ggplot2::geom_tile(
        fill = "white",
        color = "grey60",
        linewidth = 0.5
    ) +
    ggplot2::geom_text(
        ggplot2::aes(
            label = displayed_value
        ),
        size = 3
    ) +
    ggplot2::coord_equal() +
    ggplot2::labs(
        title = "Pairwise feature covariance matrix",
        subtitle = "Transformed training predictors",
        x = NULL,
        y = NULL
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
        panel.grid = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(
            angle = 45,
            hjust = 1,
            face = "bold"
        ),
        axis.text.y = ggplot2::element_text(
            face = "bold"
        )
    )

print(
    covariance_table_plot
)


# Save the table as PNG.

ggplot2::ggsave(
    filename = file.path(
        plot_directory,
        "7.5_feature_covariance_matrix.png"
    ),
    plot = covariance_table_plot,
    width = 9,
    height = 8,
    units = "in",
    dpi = 300,
    bg = "white"
)

# ---------------------------------------------------------
# 8. Possible anomaly
# ---------------------------------------------------------

maximum_absolute_z <- apply(
    abs(x_train),
    MARGIN = 1,
    FUN = max
)

possible_anomalies <- tibble::tibble(
    training_position = seq_len(
        nrow(x_train)
    ),
    original_row = train_id,
    maximum_absolute_z =
        maximum_absolute_z
) |>
    dplyr::arrange(
        dplyr::desc(maximum_absolute_z)
    ) |>
    dplyr::slice_head(
        n = 5
    )

knitr::kable(
    possible_anomalies,
    digits = 3,
    caption = paste(
        "Training observations with the largest",
        "absolute standardized predictor values"
    )
)

readr::write_csv(
    possible_anomalies,
    file = file.path(
        table_directory,
        "8. possible_anomalies.csv"
    )
)

most_extreme_training_position <-
    possible_anomalies$training_position[1]

most_extreme_predictor <- colnames(x_train)[
    which.max(
        abs(
            x_train[
                most_extreme_training_position,
                ,
                drop = TRUE
            ]
        )
    )
]

possible_anomaly_detail <- tibble::tibble(
    original_row =
        possible_anomalies$original_row[1],
    predictor = most_extreme_predictor,
    standardized_value = x_train[
        most_extreme_training_position,
        most_extreme_predictor
    ]
)

knitr::kable(
    possible_anomaly_detail,
    digits = 3,
    caption = "Most extreme training predictor value"
)


# ---------------------------------------------------------
# 9. Potential multicollinearity
# ---------------------------------------------------------

predictor_correlation_matrix <- stats::cor(
    x_train_transformed,
    use = "complete.obs"
)

upper_triangle_indices <- which(
    upper.tri(
        predictor_correlation_matrix
    ),
    arr.ind = TRUE
)

correlation_pairs <- tibble::tibble(
    predictor_1 = rownames(
        predictor_correlation_matrix
    )[
        upper_triangle_indices[, 1]
    ],
    predictor_2 = colnames(
        predictor_correlation_matrix
    )[
        upper_triangle_indices[, 2]
    ],
    correlation = predictor_correlation_matrix[
        upper_triangle_indices
    ]
) |>
    dplyr::mutate(
        absolute_correlation = abs(correlation)
    ) |>
    dplyr::arrange(
        dplyr::desc(absolute_correlation)
    )

strongest_correlation_pairs <- correlation_pairs |>
    dplyr::slice_head(
        n = 5
    )

knitr::kable(
    strongest_correlation_pairs,
    digits = 3,
    caption = "Strongest training predictor correlations"
)

readr::write_csv(
    strongest_correlation_pairs,
    file = file.path(
        table_directory,
        "9. strongest_predictor_correlations.csv"
    )
)

# ---------------------------------------------------------
# 10. Exploratory nonlinear-relationship check
#
# Compare:
# 1. additive OLS model;
# 2. full second-order OLS model.
#
# Training observations only.
# ---------------------------------------------------------

# Standardize the raw training predictors separately for this
# exploratory comparison. This prevents existing transformations,
# such as age -> age^2, from being squared again.

nonlinear_scaler <- fit_scaler(
    x_train_raw
)

x_train_nonlinear <- apply_scaler(
    x_train_raw,
    nonlinear_scaler
)

nonlinear_model_data <- as.data.frame(
    x_train_nonlinear
)

nonlinear_model_data$y <- y_train


# Additive reference model.

additive_ols_model <- stats::lm(
    y ~
        lcavol +
        lweight +
        age +
        lbph +
        svi +
        lcp +
        gleason +
        pgg45,
    data = nonlinear_model_data
)


# Full second-order model:
# - all main effects;
# - all pairwise interactions;
# - quadratic terms for nonbinary predictors.
#
# I(svi^2) is omitted because svi is binary, so svi^2 = svi.

quadratic_ols_model <- stats::lm(
    y ~ (
        lcavol +
        lweight +
        age +
        lbph +
        svi +
        lcp +
        gleason +
        pgg45
    )^2 +
        I(lcavol^2) +
        I(lweight^2) +
        I(age^2) +
        I(lbph^2) +
        I(lcp^2) +
        I(gleason^2) +
        I(pgg45^2),
    data = nonlinear_model_data
)


# Compare the nested models using a partial F-test.

nonlinear_model_test <- stats::anova(
    additive_ols_model,
    quadratic_ols_model
)

nonlinear_model_test_table <- nonlinear_model_test |>
    as.data.frame() |>
    tibble::rownames_to_column(
        var = "model"
    ) |>
    dplyr::mutate(
        model = c(
            "Additive OLS",
            "Second-order OLS"
        )
    )

knitr::kable(
    nonlinear_model_test_table,
    digits = 4,
    caption = paste(
        "Training-only comparison of additive",
        "and second-order OLS models"
    )
)

readr::write_csv(
    nonlinear_model_test_table,
    file = file.path(
        table_directory,
        "nonlinear_model_comparison.csv"
    )
)


# Export the second-order model coefficients.

quadratic_coefficient_table <- broom::tidy(
    quadratic_ols_model,
    conf.int = TRUE,
    conf.level = 0.95
) |>
    dplyr::mutate(
        significant_at_005 =
            !is.na(p.value) &
            p.value < 0.05
    )

knitr::kable(
    quadratic_coefficient_table,
    digits = 4,
    caption = "Second-order OLS coefficient estimates"
)

readr::write_csv(
    quadratic_coefficient_table,
    file = file.path(
        table_directory,
        "quadratic_ols_coefficients.csv"
    )
)


# Report model sizes and training fit.

nonlinear_model_summary <- dplyr::bind_rows(
    broom::glance(
        additive_ols_model
    ) |>
        dplyr::mutate(
            model = "Additive OLS",
            .before = 1
        ),
    broom::glance(
        quadratic_ols_model
    ) |>
        dplyr::mutate(
            model = "Second-order OLS",
            .before = 1
        )
) |>
    dplyr::select(
        model,
        r.squared,
        adj.r.squared,
        sigma,
        df,
        df.residual,
        AIC,
        BIC,
        nobs
    )

knitr::kable(
    nonlinear_model_summary,
    digits = 4,
    caption = "Training fit of additive and second-order models"
)

readr::write_csv(
    nonlinear_model_summary,
    file = file.path(
        table_directory,
        "nonlinear_model_summary.csv"
    )
)


# Check whether any coefficients could not be estimated.

aliased_quadratic_terms <- quadratic_coefficient_table |>
    dplyr::filter(
        is.na(estimate)
    )

if (nrow(aliased_quadratic_terms) > 0L) {
    message(
        "Some second-order coefficients are aliased ",
        "and could not be estimated."
    )

    print(
        aliased_quadratic_terms
    )
}

save_table_png <- function(
    data,
    filename,
    title,
    subtitle = NULL,
    width = 12,
    row_height = 0.35
) {
    column_order <- names(data)

    display_data <- data |>
        dplyr::mutate(
            dplyr::across(
                dplyr::where(is.numeric),
                function(x) {
                    dplyr::if_else(
                        is.na(x),
                        "",
                        formatC(
                            x,
                            format = "g",
                            digits = 4
                        )
                    )
                }
            ),
            dplyr::across(
                !dplyr::where(is.numeric),
                as.character
            )
        ) |>
        dplyr::mutate(
            table_row = dplyr::row_number()
        ) |>
        tidyr::pivot_longer(
            cols = -table_row,
            names_to = "column",
            values_to = "displayed_value"
        ) |>
        dplyr::mutate(
            column = factor(
                column,
                levels = column_order
            ),
            table_row = factor(
                table_row,
                levels = rev(
                    seq_len(nrow(data))
                )
            )
        )

    table_plot <- ggplot2::ggplot(
        display_data,
        ggplot2::aes(
            x = column,
            y = table_row
        )
    ) +
        ggplot2::geom_tile(
            fill = "white",
            color = "grey65",
            linewidth = 0.4
        ) +
        ggplot2::geom_text(
            ggplot2::aes(
                label = displayed_value
            ),
            size = 3
        ) +
        ggplot2::scale_x_discrete(
            position = "top"
        ) +
        ggplot2::labs(
            title = title,
            subtitle = subtitle,
            x = NULL,
            y = NULL
        ) +
        ggplot2::theme_minimal() +
        ggplot2::theme(
            panel.grid = ggplot2::element_blank(),
            axis.text.x = ggplot2::element_text(
                face = "bold",
                angle = 30,
                hjust = 0
            ),
            axis.text.y = ggplot2::element_blank(),
            axis.ticks = ggplot2::element_blank()
        )

    image_height <- max(
        3,
        2 + row_height * nrow(data)
    )

    ggplot2::ggsave(
        filename = file.path(
            plot_directory,
            filename
        ),
        plot = table_plot,
        width = width,
        height = image_height,
        units = "in",
        dpi = 300,
        bg = "white",
        limitsize = FALSE
    )

    print(table_plot)

    invisible(table_plot)
}


# Save the nested-model comparison.

save_table_png(
    data = nonlinear_model_test_table,
    filename = "10_nonlinear_model_comparison.png",
    title = "Additive versus second-order OLS",
    subtitle = "Training-only partial F-test",
    width = 10
)


# Save the model-level summary.

save_table_png(
    data = nonlinear_model_summary,
    filename = "10_nonlinear_model_summary.png",
    title = "Additive and second-order model summary",
    subtitle = "Training observations only",
    width = 12
)


# Save the second-order coefficient table.

save_table_png(
    data = quadratic_coefficient_table,
    filename = "10_quadratic_ols_coefficients.png",
    title = "Second-order OLS coefficients",
    subtitle = "Estimates with 95% confidence intervals",
    width = 14,
    row_height = 0.30
)

# ---------------------------------------------------------
# Values for the written interpretation
# ---------------------------------------------------------

cat(
    "\nFocused predictor:",
    focused_predictor,
    "\n"
)

cat(
    "Possible anomaly: original row",
    possible_anomaly_detail$original_row,
    "with standardized",
    possible_anomaly_detail$predictor,
    "=",
    round(
        possible_anomaly_detail$standardized_value,
        3
    ),
    "\n"
)

cat(
    "Strongest correlated predictor pair:",
    strongest_correlation_pairs$predictor_1[1],
    "and",
    strongest_correlation_pairs$predictor_2[1],
    "with correlation",
    round(
        strongest_correlation_pairs$correlation[1],
        3
    ),
    "\n"
)
```

*Interpret multicollinearity, one anomaly, and the question regularization will address.*

# Problem 2. OLS, ridge, and lasso

## 2A. OLS baseline

```{r ols, eval=FALSE}
# Fit OLS on x_train and y_train. Do not access y_test here.
```

## 2B. Ridge

```{r ridge, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 0, foldid = foldid).
# Report lambda.min, lambda.1se, paths, coefficients, and CV evidence.
```

## 2C. Lasso

```{r lasso, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 1, foldid = foldid).
# Report lambda.min, lambda.1se, paths, and nonzero predictors.
```

## 2D. Training-based comparison and locked core model

```{r core-comparison, eval=FALSE}
# Build a comparison from training/CV quantities only.
```

**Core model nominated before holdout evaluation:** [method and reason]

## 2E. Perturbation or stability check

**Prediction before running the check:** [prediction]

```{r perturbation, eval=FALSE}
# Implement one allowed training-only stability check.
```

**Observed result and explanation:** [result]

# Problem 3. Mathematical mechanisms

## 3A. Ridge and conditioning

*Write the ridge derivation. Define dimensions and explain uniqueness.*

```{r ridge-conditioning, eval=FALSE}
# Use x_train and the fitted ridge lambda.min.
# safe_condition_numbers(x_train, lambda = ...)
```

## 3B. Lasso optimality and soft thresholding

*Derive the zero/nonzero optimality conditions and the orthonormal-design formula.*

## 3C. Connect algebra to evidence

*Connect one specific numerical result to Problem 3A or 3B.*

# Problem 4. Elastic net and a neural-feature bridge

## 4A. Research question and source map

**Research direction:** [weight decay / sparse weights / pruning / dropout]

| Claim | Exact primary source and locator | What it establishes | What it does not establish |
|---|---|---|---|
| [Claim] | [Citation, section/page/equation] | [Support] | [Boundary] |

## 4B. Elastic net on original predictors

```{r original-elastic-net, eval=FALSE}
alpha_grid <- c(0.10, 0.25, 0.50, 0.75, 0.90)

# Fit one cv.glmnet object per alpha with the same foldid.
# Select alpha and lambda using training CV only.
```

## 4C. Fixed ReLU features

```{r random-relu-features, eval=FALSE}
relu <- function(z) pmax(z, 0)
M <- 200L

# TODO:
# 1. Generate A and bias once from seeds$features using the stated distributions.
# 2. Apply the same A and bias to x_train and x_test.
# 3. Fit a scaler to H_train only and apply it to H_test.
# 4. Fit ridge, lasso, and elastic net on H_train with the shared foldid.
```

*State which weights are fixed, which are learned, and how active features are counted.*

## 4D. Locked declaration and final holdout evaluation

> **Locked before viewing `y_test_locked`:** [preprocessing, candidate specifications, selected tuning values, feature seed, nominated deployment model, metrics]

```{r final-holdout-evaluation, eval=FALSE}
# This must be the first analysis chunk that transforms and uses y_test_locked.
# Use y_test <- transform_response(y_test_locked) here, then evaluate every
# already-fixed candidate on the sqrt(PSA) response scale.
# Generate all fixed candidate predictions and one RMSE/MAE table here.
```

*State whether the holdout evidence supports the predeclared choice. Do not retune.*

## 4E. Deep-learning connection and limitations

*Distinguish penalty, weight decay, output-weight sparsity, pruning, dropout, and a trained deep network. Cite the sources from 4A and discuss at least two limitations.*

# Problem 5. Report quality and reproducibility

## 5A. Reproduction record

```text
1. Open a clean R session.
2. [Exact commands and folder layout.]
3. Render Group02_ProjectQuiz1.Rmd.
```

## 5B. Recommendation and limitations

*Give the final recommendation, distinguish prediction from interpretation, and state limitations.*

## 5C. AI verification record

| AI-assisted item | How it was checked | Evidence | Modification or rejection |
|---|---|---|---|
| [Item] | [Method] | [Test/source/output] | [Decision] |

# Preflight and session information {-}

```{r preflight, eval=FALSE}
# Add every object required to reproduce the final table and figures.
required_objects <- c(
  "train_id", "test_id", "x_train", "x_test", "y_train",
  "y_test_locked", "transform_response", "foldid"
)
missing_objects <- setdiff(required_objects, ls())
if (length(missing_objects)) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}
stopifnot(length(intersect(train_id, test_id)) == 0L)
```

```{r session-info}
sessionInfo()
```
