---
title: "ASESII 24A02 Project-Based Quiz 1"
subtitle: "Ridge, Lasso, and Regularization for Neural Features"
author: "Group 02: Nguyễn Thanh An - 24125024, Nguyễn Chí Bảo - 24125025, Trần Nguyễn Khánh Bình - 24125026, Phạm Văn Trang - 24125046, Quách Thiên Lạc - 24125092"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    fig_crop: false
  html_document:
    toc: true
    toc_depth: 3
bibliography: references.bib
link-citations: true
params:
  group_number: 2
fontsize: 11pt
geometry: margin=1in
---

<!--
Copy this template into your group submission folder, rename it
Group02_ProjectQuiz1.Rmd, place references.bib beside it, and set the group
number parameter before rendering. Remove instructional comments from the
final report.
-->

# Authorship and contribution statement {-}

We confirm that every group member can explain the submitted analysis. AI use is reported in `Group02_AI_Log.csv`, and the group checked every AI-assisted item used in this report.

| Student | ID | Main contributions | Verification performed |
|---|---|---|---|
| Nguyễn Thanh An        | 24125024 |---|---|
| Nguyễn Chí Bảo         | 24125025 |---|---|
| Trần Nguyễn Khánh Bình | 24125026 |---|---|
| Phạm Văn Trang         | 24125046 |---|---|
| Quách Thiên Lạc        | 24125092 |---|---|


# Abstract {-}

*Maximum 150 words: prediction problem, methods, main evidence, and recommendation.*

# Reproducible setup {-}

```{r setup, message=FALSE, warning=FALSE}
required_packages <- c("tidyverse", "glmnet", "broom", "knitr")
missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages)) {
  stop("Install the required packages before rendering: ",
       paste(missing_packages, collapse = ", "))
}

library(tidyverse)
library(glmnet)
library(broom)
library(knitr)

# hardcode group number bc somehow the object group_number in param is not found
group_number <- 2

# group_number <- as.integer(params$group_number)
# if (length(group_number) != 1L || is.na(group_number) || group_number < 1L) {
#   stop("Set params$group_number to your assigned positive group number.")
# }

seeds <- list(
  split = 240200L + group_number,
  cv = 240300L + group_number,
  features = 240400L + group_number
)
```

## Shared helper functions {-}

```{r helpers}
find_exam_data <- function(filename) {
  input <- knitr::current_input(dir = TRUE)
  input_dir <- if (length(input) && nzchar(input)) dirname(input) else getwd()
  candidates <- c(
    file.path(input_dir, "data", filename),
    file.path(input_dir, "..", "data", filename),
    file.path(getwd(), "data", filename)
  )
  existing <- candidates[file.exists(candidates)]
  if (!length(existing)) stop("Cannot locate ", filename, " by a relative path.")
  hits <- unique(normalizePath(existing, winslash = "/", mustWork = TRUE))
  if (length(hits) > 1L && length(unique(unname(tools::md5sum(hits)))) > 1L) {
    stop("Multiple nonidentical copies of ", filename, " were found.")
  }
  hits[[1L]]
}

split_rows <- function(n, train_prop = 0.80, seed) {
  stopifnot(n >= 10L, train_prop > 0, train_prop < 1)
  set.seed(seed)
  train <- sort(sample.int(n, floor(train_prop * n), replace = FALSE))
  list(train = train, test = setdiff(seq_len(n), train))
}

fit_scaler <- function(x, tol = 1e-12) {
  x <- as.matrix(x)
  center <- colMeans(x)
  centered <- sweep(x, 2L, center, "-")
  scale <- sqrt(colMeans(centered^2))
  keep <- is.finite(scale) & scale > tol
  if (!any(keep)) stop("No nonconstant training predictor remains.")
  list(
    center = center[keep],
    scale = scale[keep],
    keep = keep,
    dropped = colnames(x)[!keep]
  )
}

apply_scaler <- function(x, prep) {
  x <- as.matrix(x)[, prep$keep, drop = FALSE]
  x <- sweep(x, 2L, prep$center, "-")
  sweep(x, 2L, prep$scale, "/")
}

make_foldid <- function(n, k = 5L, seed) {
  if (k < 3L || k > n) stop("Require 3 <= k <= number of training rows.")
  set.seed(seed)
  sample(rep(seq_len(k), length.out = n))
}

fit_cv_glmnet <- function(x, y, alpha, foldid) {
  glmnet::cv.glmnet(
    x = x,
    y = y,
    family = "gaussian",
    alpha = alpha,
    foldid = foldid,
    standardize = FALSE,
    intercept = TRUE,
    type.measure = "mse"
  )
}

score_regression <- function(y, prediction) {
  y <- as.numeric(y)
  prediction <- as.numeric(prediction)
  if (length(y) != length(prediction) || any(!is.finite(c(y, prediction)))) {
    stop("Metrics require equal-length finite vectors.")
  }
  c(RMSE = sqrt(mean((y - prediction)^2)),
    MAE = mean(abs(y - prediction)))
}

count_nonzero <- function(fit, s = "lambda.min", tol = 1e-8) {
  b <- as.matrix(stats::coef(fit, s = s))
  slopes <- b[rownames(b) != "(Intercept)", 1L]
  sum(abs(slopes) > tol)
}

safe_condition_numbers <- function(x, lambda, tol = 1e-10) {
  eigenvalues <- eigen(
    crossprod(x) / nrow(x),
    symmetric = TRUE,
    only.values = TRUE
  )$values
  largest <- max(eigenvalues)
  smallest <- max(min(eigenvalues), 0)
  cutoff <- tol * max(1, largest)
  c(
    gram = if (smallest <= cutoff) Inf else largest / smallest,
    ridge = (largest + lambda) / (smallest + lambda)
  )
}
```

# Problem 1. Prediction design and leakage control

## 1A. Target and predictors

```{r data-configuration}
config <- if (group_number %% 2L == 1L) {
  list(
    file = "fat.csv",
    response = "brozek",
    excluded = c("brozek", "siri", "density", "free")
  )
} else {
  list(
    file = "prostate.csv",
    response = "lpsa",
    excluded = "lpsa"
  )
}

data_raw <- read.csv(find_exam_data(config$file), check.names = FALSE)
predictors <- setdiff(names(data_raw), config$excluded)
analysis_data <- data_raw[, c(config$response, predictors), drop = FALSE]

if (anyNA(analysis_data)) stop("Declare a training-only missing-value plan.")
if (!all(vapply(analysis_data, is.numeric, logical(1)))) {
  stop("The assigned response and predictors must be numeric.")
}

# csv output for preflight check
cat("File:", config$file, "\n")
cat("Response:", config$response, "\n")
cat("Rows:", nrow(data_raw), "\n")
cat("Columns:", ncol(data_raw), "\n")

# basic checks 
integrity_summary <- tibble::tibble(
    variable = names(data_raw),
    type = vapply(
        data_raw,
        function(x) class(x)[1],
        character(1)
    ),
    missing = vapply(
        data_raw,
        function(x) sum(is.na(x)),
        integer(1)
    ),
    unique_values = vapply(
        data_raw,
        function(x) dplyr::n_distinct(
            x,
            na.rm = TRUE
        ),
        integer(1)
    )
)

knitr::kable(
    integrity_summary,
    caption = "Basic CSV integrity checks"
)

# find missing values
missing_response_count <- sum(
    is.na(analysis_data[[config$response]])
)

cat(
    "Rows with missing responses:",
    missing_response_count,
    "\n"
)

analysis_data <- analysis_data |>
    filter(
        !is.na(.data[[config$response]])
    )
```

*Define the prediction task and justify every exclusion.*

## 1B. Split and train-only preprocessing

```{r split-and-scale}
# split the data 80/20 into training  and test sets, using the group-specific seed

split <- split_rows(
    n = nrow(analysis_data),
    train_prop = 0.80,
    seed = seeds$split
)

train_id <- split$train
test_id <- split$test

stopifnot(
    length(intersect(train_id, test_id)) == 0L
)


# make raw training and test predictor matrices and response vectors, without any preprocessing yet

x_train_raw <- analysis_data[
    train_id,
    predictors,
    drop = FALSE
]

x_test_raw <- analysis_data[
    test_id,
    predictors,
    drop = FALSE
]

y_train <- analysis_data[
    train_id,
    config$response,
    drop = TRUE
]

# dont touch this 
y_test_locked <- analysis_data[
    test_id,
    config$response,
    drop = TRUE
]


# start calc. the skewness of each continuous predictor in the training set, to identify which ones may need transformation
sample_skewness <- function(x) {
  # check if x is numeric 
    x <- x[is.finite(x)]
    n <- length(x)

    if (n < 3L) {
        return(NA_real_)
    }
    # standardize x
    standard_deviation <- stats::sd(x)

    if (
        !is.finite(standard_deviation) ||
        standard_deviation == 0
    ) {
        return(NA_real_)
    }

    standardized <- (
        x - mean(x)
    ) / standard_deviation

    n / (
        (n - 1) * (n - 2)
    ) * sum(standardized^3)
}


# find continuous predictors: numeric and more than 5 unique values
continuous_predictors <- predictors[
    vapply(
        x_train_raw,
        function(x) {
            is.numeric(x) &&
                dplyr::n_distinct(
                    x,
                    na.rm = TRUE
                ) > 5L
        },
        logical(1)
    )
]

noncontinuous_predictors <- setdiff(
    predictors,
    continuous_predictors
)


# start calc. skewness to see if any continuous predictors are strongly skewed and may need trans.

# function to safely compute the minimum of a vector, returning NA if no finite values exist
safe_minimum <- function(x) {
    x <- x[is.finite(x)]

    if (length(x) == 0L) {
        return(NA_real_)
    }

    min(x)
}

  # make a skewness table for continuous predictors, including minimum, skewness, absolute skewness, and assessment of skewness level (important for analysis later on)
skewness_table <- tibble::tibble(
    variable = continuous_predictors,
    minimum = vapply(
        x_train_raw[continuous_predictors],
        safe_minimum,
        numeric(1)
    ),
    skewness = vapply(
        x_train_raw[continuous_predictors],
        sample_skewness,
        numeric(1)
    )
) |>
# here we define the absolute skewness and assessment of skewness level based on the absolute skewness value
    dplyr::mutate(
        absolute_skewness = abs(skewness),
        assessment = dplyr::case_when(
            is.na(absolute_skewness) ~
                "Unavailable",
            absolute_skewness < 0.5 ~
                "Approximately symmetric",
            absolute_skewness < 1 ~
                "Moderately skewed",
            TRUE ~
                "Strongly skewed"
        )
    ) |>
    dplyr::arrange(
        dplyr::desc(absolute_skewness)
    )

knitr::kable(
    skewness_table,
    digits = 3,
    caption = "Training-set predictor skewness"
)


# find the predictors that are already logarithmic to avoid transforming them again
already_logged <- if (config$response == "lpsa") {
    intersect(
        predictors,
        c(
            "lcavol",
            "lweight",
            "lbph",
            "lcp"
        )
    )
} else {
    character(0)
}


# ---------------------------------------------------------
# Select predictors for log1p transformation
#
# Rule:
# - continuous;
# - skewness >= 1;
# - nonnegative;
# - not already logarithmic.
# ---------------------------------------------------------

log1p_variables <- skewness_table |>
    dplyr::filter(
        !is.na(skewness),
        skewness >= 1,
        !is.na(minimum),
        minimum >= 0,
        !variable %in% already_logged
    ) |>
    dplyr::pull(variable)

cat(
    "Predictors selected for log1p transformation:\n"
)

if (length(log1p_variables) == 0L) {
    cat("None\n")
} else {
    print(log1p_variables)
}


# validate that the test set does not contain negative values for any log1p variables, which would be invalid for the transformation

if (length(log1p_variables) > 0L) {
    invalid_test_domains <- vapply(
        x_test_raw[log1p_variables],
        function(x) {
            any(
                x < 0,
                na.rm = TRUE
            )
        },
        logical(1)
    )

    if (any(invalid_test_domains)) {
        stop(
            "Negative test values found for log1p variables: ",
            paste(
                names(invalid_test_domains)[
                    invalid_test_domains
                ],
                collapse = ", "
            )
        )
    }
}


# create transformed training and test predictor matrices, applying log1p to the selected variables

x_train_transformed <- x_train_raw
x_test_transformed <- x_test_raw

for (variable in log1p_variables) {
    x_train_transformed[[variable]] <- log1p(
        x_train_raw[[variable]]
    )

    x_test_transformed[[variable]] <- log1p(
        x_test_raw[[variable]]
    )
}


# compare the skewness before and after transformation for the selected variables, to assess the effectiveness of the log1p transformation (also important for analysis later on)
if (length(log1p_variables) > 0L) {
    transformation_comparison <- tibble::tibble(
        variable = log1p_variables,
        original_skewness = vapply(
            x_train_raw[log1p_variables],
            sample_skewness,
            numeric(1)
        ),
        transformed_skewness = vapply(
            x_train_transformed[log1p_variables],
            sample_skewness,
            numeric(1)
        )
    ) |>
        dplyr::mutate(
            absolute_reduction =
                abs(original_skewness) -
                abs(transformed_skewness)
        )

    print(
        knitr::kable(
            transformation_comparison,
            digits = 3,
            caption = paste(
                "Training-set skewness before",
                "and after transformation"
            )
        )
    )
}


# calc. the median of each predictor in the training set, to use for imputation of missing values in both training and test sets (just to be extra sure - but I checked the data after and looks like this can be dropped)
training_medians <- vapply(
    x_train_transformed,
    function(x) {
        stats::median(
            x,
            na.rm = TRUE
        )
    },
    numeric(1)
)

invalid_medians <- !is.finite(training_medians)

if (any(invalid_medians)) {
    stop(
        "No usable training values for predictors: ",
        paste(
            names(training_medians)[invalid_medians],
            collapse = ", "
        )
    )
}

impute_with_medians <- function(data, medians) {
    result <- data

    for (variable in names(medians)) {
        missing_positions <- is.na(
            result[[variable]]
        )

        result[[variable]][missing_positions] <-
            medians[[variable]]
    }

    result
}

x_train_imputed <- impute_with_medians(
    x_train_transformed,
    training_medians
)

x_test_imputed <- impute_with_medians(
    x_test_transformed,
    training_medians
)

stopifnot(
    !anyNA(x_train_imputed),
    !anyNA(x_test_imputed)
)


# fit the scaler on the training set and apply it to both training and test sets

x_prep <- fit_scaler(
    x_train_imputed
)

x_train <- apply_scaler(
    x_train_imputed,
    x_prep
)

x_test <- apply_scaler(
    x_test_imputed,
    x_prep
)


# create foldid for cross-validation, using the group-specific seed (same as original code, but now we are using the transformed and scaled x_train from the previous steps)
foldid <- make_foldid(
    n = nrow(x_train),
    k = 5L,
    seed = seeds$cv
)

cat("\nFold sizes:\n")
print(table(foldid))


# summarize the final transformation rules for each predictor, indicating whether log1p was applied or not

transformation_record <- tibble::tibble(
    variable = predictors,
    transformation = dplyr::if_else(
        predictors %in% log1p_variables,
        "log1p(x)",
        "none"
    )
)

knitr::kable(
    transformation_record,
    caption = "Final training-derived transformation rules"
)

# export the tibble frames to a csv file
readr::write_csv(
  skewness_table,
  file = "../data/tables/skewness_table.csv"
)
readr::write_csv(
    transformation_record,
    file = "../data/tables/transformation_record.csv"
)
```

*Report seeds, dimensions, scaling, folds, and leakage safeguards.*

## 1C. Training-data exploration

```{r training-eda, fig.width=7, fig.height=5}
# Use analysis_data[train_id, ] only. Add focused plots and summaries.

# add skewness plots from 1B for the selected log1p variables, showing the distribution before and after transformation

# 1. training response distribution
response_plot_data <- tibble::tibble(
    response = y_train
)

ggplot2::ggplot(
    response_plot_data,
    ggplot2::aes(x = response)
) +
    ggplot2::geom_histogram(
        bins = 20,
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::labs(
        title = paste(
            "Training distribution of",
            config$response
        ),
        x = config$response,
        y = "Count"
    ) +
    ggplot2::theme_minimal()

  ggplot2::ggsave(
    filename = paste0(
        "../data/training_distribution_",
        config$response,
        ".png"
    ),
    width = 7,
    height = 5,
    units = "in",
    dpi = 300
  )

# 2. plot discrete predictors as bar plots, showing the count of each category in the training set
discrete_predictors <- intersect(
    c(
        "svi",
        "gleason"
    ),
    predictors
)

discrete_plot_data <- x_train_raw |>
    dplyr::select(
        dplyr::all_of(discrete_predictors)
    ) |>
    tidyr::pivot_longer(
        cols = dplyr::everything(),
        names_to = "variable",
        values_to = "value"
    ) |>
    dplyr::count(
        variable,
        value,
        name = "count"
    )

discrete_predictor_plot <- ggplot2::ggplot(
    discrete_plot_data,
    ggplot2::aes(
        x = factor(value),
        y = count
    )
) +
    ggplot2::geom_col(
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::facet_wrap(
        ~ variable,
        scales = "free_x"
    ) +
    ggplot2::labs(
        title = "Discrete predictor distributions",
        subtitle = "Training observations only",
        x = "Value",
        y = "Count"
    ) +
    ggplot2::theme_minimal()

discrete_predictor_plot

  ggplot2::ggsave(
    filename = "../data/original_discrete_predictor_distributions.png",
    width = 7,
    height = 5,
    units = "in",
    dpi = 300
  )

# 3. original predictor distributions for continuous predictors, before any transformation

original_plot_data <- x_train_raw |>
    dplyr::select(
        dplyr::all_of(continuous_predictors)
    ) |>
    tidyr::pivot_longer(
        cols = dplyr::everything(),
        names_to = "variable",
        values_to = "value"
    )

ggplot2::ggplot(
    original_plot_data,
    ggplot2::aes(x = value)
) +
    ggplot2::geom_histogram(
        bins = 20,
        fill = "steelblue",
        color = "white"
    ) +
    ggplot2::facet_wrap(
        ~ variable,
        scales = "free",
        ncol = 3
    ) +
    ggplot2::labs(
        title = "Original predictor distributions",
        subtitle = "Training observations only",
        x = NULL,
        y = "Count"
    ) +
    ggplot2::theme_minimal()

    ggplot2::ggsave(
    filename = "../data/original_continuous_predictor_distributions.png",
    width = 7,
    height = 5,
    units = "in",
    dpi = 300
)
# 4. comparison of distributions before and after log1p transformation for the selected variables

if (length(log1p_variables) > 0L) {
    before_plot_data <- x_train_raw |>
        dplyr::select(
            dplyr::all_of(log1p_variables)
        ) |>
        tidyr::pivot_longer(
            cols = dplyr::everything(),
            names_to = "variable",
            values_to = "value"
        ) |>
        dplyr::mutate(
            version = "Original"
        )

    after_plot_data <- x_train_transformed |>
        dplyr::select(
            dplyr::all_of(log1p_variables)
        ) |>
        tidyr::pivot_longer(
            cols = dplyr::everything(),
            names_to = "variable",
            values_to = "value"
        ) |>
        dplyr::mutate(
            version = "log1p transformed"
        )

    comparison_plot_data <- dplyr::bind_rows(
        before_plot_data,
        after_plot_data
    )

    print(
        ggplot2::ggplot(
            comparison_plot_data,
            ggplot2::aes(x = value)
        ) +
            ggplot2::geom_histogram(
                bins = 20,
                fill = "steelblue",
                color = "white"
            ) +
            ggplot2::facet_grid(
                variable ~ version,
                scales = "free"
            ) +
            ggplot2::labs(
                title = paste(
                    "Predictor distributions before",
                    "and after transformation"
                ),
                subtitle = "Training observations only",
                x = NULL,
                y = "Count"
            ) +
            ggplot2::theme_minimal()
    )
}

  ggplot2::ggsave(
    filename = "../data/after_transformation_comparison.png",
    width = 7,
    height = 5,
    units = "in",
    dpi = 300
  )
```

*Interpret multicollinearity, one anomaly, and the question regularization will address.*

# Problem 2. OLS, ridge, and lasso

## 2A. OLS baseline

```{r ols, eval=FALSE}
# Fit OLS on x_train and y_train. Do not access y_test here.
```

## 2B. Ridge

```{r ridge, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 0, foldid = foldid).
# Report lambda.min, lambda.1se, paths, coefficients, and CV evidence.
```

## 2C. Lasso

```{r lasso, eval=FALSE}
# Use fit_cv_glmnet(..., alpha = 1, foldid = foldid).
# Report lambda.min, lambda.1se, paths, and nonzero predictors.
```

## 2D. Training-based comparison and locked core model

```{r core-comparison, eval=FALSE}
# Build a comparison from training/CV quantities only.
```

**Core model nominated before holdout evaluation:** [method and reason]

## 2E. Perturbation or stability check

**Prediction before running the check:** [prediction]

```{r perturbation, eval=FALSE}
# Implement one allowed training-only stability check.
```

**Observed result and explanation:** [result]

# Problem 3. Mathematical mechanisms

## 3A. Ridge and conditioning

*Write the ridge derivation. Define dimensions and explain uniqueness.*

```{r ridge-conditioning, eval=FALSE}
# Use x_train and the fitted ridge lambda.min.
# safe_condition_numbers(x_train, lambda = ...)
```

## 3B. Lasso optimality and soft thresholding

*Derive the zero/nonzero optimality conditions and the orthonormal-design formula.*

## 3C. Connect algebra to evidence

*Connect one specific numerical result to Problem 3A or 3B.*

# Problem 4. Elastic net and a neural-feature bridge

## 4A. Research question and source map

**Research direction:** [weight decay / sparse weights / pruning / dropout]

| Claim | Exact primary source and locator | What it establishes | What it does not establish |
|---|---|---|---|
| [Claim] | [Citation, section/page/equation] | [Support] | [Boundary] |

## 4B. Elastic net on original predictors

```{r original-elastic-net, eval=FALSE}
alpha_grid <- c(0.10, 0.25, 0.50, 0.75, 0.90)

# Fit one cv.glmnet object per alpha with the same foldid.
# Select alpha and lambda using training CV only.
```

## 4C. Fixed ReLU features

```{r random-relu-features, eval=FALSE}
relu <- function(z) pmax(z, 0)
M <- 200L

# TODO:
# 1. Generate A and bias once from seeds$features using the stated distributions.
# 2. Apply the same A and bias to x_train and x_test.
# 3. Fit a scaler to H_train only and apply it to H_test.
# 4. Fit ridge, lasso, and elastic net on H_train with the shared foldid.
```

*State which weights are fixed, which are learned, and how active features are counted.*

## 4D. Locked declaration and final holdout evaluation

> **Locked before viewing `y_test`:** [preprocessing, candidate specifications, selected tuning values, feature seed, nominated deployment model, metrics]

```{r final-holdout-evaluation, eval=FALSE}
# This must be the first analysis chunk that uses y_test for evaluation.
# Generate all fixed candidate predictions and one RMSE/MAE table here.
```

*State whether the holdout evidence supports the predeclared choice. Do not retune.*

## 4E. Deep-learning connection and limitations

*Distinguish penalty, weight decay, output-weight sparsity, pruning, dropout, and a trained deep network. Cite the sources from 4A and discuss at least two limitations.*

# Problem 5. Report quality and reproducibility

## 5A. Reproduction record

```text
1. Open a clean R session.
2. [Exact commands and folder layout.]
3. Render Group02_ProjectQuiz1.Rmd.
```

## 5B. Recommendation and limitations

*Give the final recommendation, distinguish prediction from interpretation, and state limitations.*

## 5C. AI verification record

| AI-assisted item | How it was checked | Evidence | Modification or rejection |
|---|---|---|---|
| [Item] | [Method] | [Test/source/output] | [Decision] |

# Preflight and session information {-}

```{r preflight, eval=FALSE}
# Add every object required to reproduce the final table and figures.
required_objects <- c(
  "train_id", "test_id", "x_train", "x_test", "y_train", "y_test", "foldid"
)
missing_objects <- setdiff(required_objects, ls())
if (length(missing_objects)) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}
stopifnot(length(intersect(train_id, test_id)) == 0L)
```

```{r session-info}
sessionInfo()
```
